\chapter{Flattening Stratification of a Coherent Sheaf}
\label{chap:Picard_FlatteningStratification}
% archon:covers AlgebraicJacobian/Picard/FlatteningStratification.lean

% STRATEGY NOTE (iter-174). This chapter is the A.2.a sub-row of Route~A
% (positive-genus arm of nonempty_jacobianWitness). It packages the
% Nitsure~\S4 flattening-stratification theorem, the technical heart that
% turns the absolute Quot-functor representability problem on a projective
% morphism \(X \to S\) with coherent input \(\F\) into a finite collection of
% locally-closed strata over which \(\F\) becomes flat. Mathlib (master
% \texttt{b80f227}) has neither a coherent-sheaf flatness predicate over a
% scheme base nor the flattening-stratification construction; this chapter
% is the prover-ready specification for the new file
% \texttt{AlgebraicJacobian/Picard/FlatteningStratification.lean} which a
% subsequent iter (file-skeleton lane) will scaffold.
%
% This chapter is independently parallel-startable: it does not gate on the
% Route~A.1.a relative-Spec chapter \cref{chap:Picard_RelativeSpec}.

\section{Setup and motivation}
\label{sec:flatstrat_setup}

The construction of the Quot scheme \(\Quot_{\F/X/S}^{\Phi,L}\) in Nitsure's exposition
of Grothendieck's existence theorem (\cite{nitsure-hilbert-quot}, \S5) proceeds by
embedding the functor of points of \(\Quot\) into a Grassmannian via Castelnuovo--Mumford
regularity, then cutting out the locus of \emph{quotients with flat fibres of the
prescribed Hilbert polynomial} by a closed-subscheme condition.  Cutting out this
locus uses, as a primary technical input, the existence of the
\emph{flattening stratification} of a coherent sheaf \(\F\) on a projective
\(S\)-scheme: a finite collection of locally-closed subschemes of \(S\) on each of which
the restriction of \(\F\) becomes flat.

This chapter packages the Nitsure~\S4 flattening-stratification theorem as the
foundational A.2.a building block. The Quot-scheme construction itself (A.2.b),
the relative Picard functor (A.1.c), and the assembly of the Picard scheme
representability theorem on a smooth proper curve (A.2.c) live in separate sibling
chapters.

We work throughout with a noetherian base scheme \(S\), a projective morphism
\(\pi : X \to S\), and a coherent \(\OO_X\)-module \(\F\). The conclusion of Nitsure's
theorem covers an arbitrary noetherian base with a projective \(X \to S\), which
specialises directly to the Route~A consumer setting where \(S = T\) is an arbitrary
\(k\)-scheme, \(X = C \times_k T\) for \(C\) a smooth proper curve over the base field
\(k\), and \(\F\) is a coherent quotient sheaf produced by the Grassmannian embedding
of the Quot functor (cf.\ \ below).

\section{Coherent flatness over a base}
\label{sec:coh_flat}

\begin{definition}

\leanok
[Coherent sheaf flat over the base]
  \label{def:coherent_sheaf_flat}
  \lean{AlgebraicGeometry.Scheme.CoherentSheafFlat}
  Let \(S\) be a noetherian scheme, \(f : X \to S\) a morphism of finite type, and
  \(\F\) a coherent \(\OO_X\)-module.  We say \(\F\) is \emph{flat over \(S\)}
  (equivalently, \(\F\) is \emph{\(f\)-flat}, or \emph{\(S\)-flat}) if for every
  point \(x \in X\) the stalk \(\F_x\) is a flat \(\OO_{S, f(x)}\)-module, where the
  \(\OO_{S, f(x)}\)-module structure is given by the composite
  \(\OO_{S, f(x)} \to \OO_{X, x} \to \End(\F_x)\).
  \textit{Source: cf.\ [Stacks Project], tag 00HB (flat modules); [Nitsure], \S4
  passim.}
  Equivalently (after pulling back to an affine open \(U = \Spec A \subseteq S\) and
  an affine open \(V = \Spec B \subseteq f^{-1}(U)\) with \(\F|_V = \widetilde{M}\)
  for \(M\) a finite \(B\)-module), \(\F|_V\) is \(\OO_U\)-flat in the sense that \(M\) is a
  flat \(A\)-module (with the \(A\)-module structure coming from the ring map
  \(A \to B\)).
\end{definition}

\begin{remark}
\label{rem:coh_flat_pullback}
  The condition is stable under base change: for any morphism \(g : S' \to S\),
  pulling back along the cartesian square
  \[
    \begin{tikzcd}
      X' \ar[r, "g'"] \ar[d, "f'"'] & X \ar[d, "f"] \\
      S' \ar[r, "g"]                & S
    \end{tikzcd}
  \]
  yields a coherent \(\OO_{X'}\)-module \(\F' := (g')^* \F\) which is \(S'\)-flat
  whenever \(\F\) is \(S\)-flat.  This is the routine cancellation of tensor-product
  associativity at the level of stalks. The converse fails in general.
\end{remark}

\section{Generic flatness}
\label{sec:generic_flatness}

The construction of the flattening stratification proceeds by Noetherian
induction on \(S\). The induction step relies on the following classical
\emph{generic flatness} theorem of Grothendieck: over a noetherian integral
base, a coherent sheaf is flat over a non-empty open subscheme. Algebraically,
this is the assertion that a finitely generated module over a finite-type
algebra over a noetherian domain is generically free after inverting a single
non-zero element of the base ring.

\subsection{The dévissage layer}
\label{sec:generic_freeness_devissage}

The proof of the algebraic generic-freeness theorem has two layers: a
\emph{dévissage} reducing an arbitrary finite module over the (noetherian)
finite-type algebra \(B\) to the case of a quotient domain \(B/\pp\), and a
\emph{normalisation core} handling that case. We first develop the dévissage
layer as standalone module theory.

\begin{definition}
\leanok
[Generically free module]
  \label{def:generically_free}
  \lean{AlgebraicGeometry.GenericFreeness.GenericallyFree}
  Let \(A\) be a commutative ring and \(M\) an \(A\)-module.  We say \(M\) is
  \emph{generically free over \(A\)} if there exists \(f \in A\), \(f \ne 0\),
  such that the localisation \(M_f\) is a free \(A_f\)-module.
\end{definition}

\begin{lemma}
\leanok
[Transport of generic freeness along further localisation]
  \label{lem:generically_free_transport}
  \lean{AlgebraicGeometry.GenericFreeness.free_localizedModule_powers_mul}
  Let \(A\) be a commutative ring, \(M\) an \(A\)-module, and \(f, g \in A\).
  If \(M_f\) is a free \(A_f\)-module, then \(M_{fg}\) is a free
  \(A_{fg}\)-module.
\end{lemma}

\begin{proof}
\leanok
  Localisation is base change: \(M_h \cong A_h \otimes_A M\) as
  \(A_h\)-modules for any \(h \in A\).  Since \(fg\) maps to a unit multiple
  of \(g\) in \(A_{fg}\), the ring \(A_{fg}\) is canonically an
  \(A_f\)-algebra, and base-change associativity gives
  \[
    M_{fg} \;\cong\; A_{fg} \otimes_A M
    \;\cong\; A_{fg} \otimes_{A_f} \bigl(A_f \otimes_A M\bigr)
    \;\cong\; A_{fg} \otimes_{A_f} M_f .
  \]
  The base change of a free module along any ring map is free (a basis of
  \(M_f\) over \(A_f\) tensors to a basis over \(A_{fg}\)), so \(M_{fg}\) is
  free over \(A_{fg}\).
\end{proof}

\begin{lemma}
\leanok
[Generic freeness is invariant under isomorphism]
  \label{lem:generically_free_of_equiv}
  \lean{AlgebraicGeometry.GenericFreeness.GenericallyFree.of_linearEquiv}
  \uses{def:generically_free}
  Let \(A\) be a commutative ring and \(M \cong N\) isomorphic \(A\)-modules.
  If \(M\) is generically free over \(A\), so is \(N\).
\end{lemma}

\begin{proof}
\leanok
  An isomorphism \(e : M \to N\) localises to an isomorphism
  \(M_f \to N_f\) of \(A_f\)-modules for the witness \(f\), and freeness
  transports along it.
\end{proof}

\begin{lemma}
\leanok
[Splicing: generic freeness is stable under extensions]
  \label{lem:generically_free_splice}
  \lean{AlgebraicGeometry.GenericFreeness.genericallyFree_of_exact}
  \uses{def:generically_free}
  \source{nitsure-hilbert-quot}
  Let \(A\) be a domain and \(0 \to M_1 \to M_2 \to M_3 \to 0\) a short exact
  sequence of \(A\)-modules.  If \(M_1\) and \(M_3\) are generically free over
  \(A\), then so is \(M_2\).
\end{lemma}

\begin{proof}
\leanok
  \uses{lem:generically_free_transport}
  Let \(f_1, f_3 \ne 0\) be witnesses for \(M_1\), \(M_3\) and set
  \(f := f_1 f_3 \ne 0\) (here \(A\) is a domain).  Localisation at \(f\) is
  exact, so \(0 \to (M_1)_f \to (M_2)_f \to (M_3)_f \to 0\) is a short exact
  sequence of \(A_f\)-modules.  By
  \cref{lem:generically_free_transport}, \((M_1)_f\) and \((M_3)_f\) are free
  \(A_f\)-modules.  A short exact sequence with free (hence projective)
  quotient splits, so \((M_2)_f \cong (M_1)_f \oplus (M_3)_f\) is free.
\end{proof}

\begin{lemma}
\leanok
[Generic freeness of the zero module]
  \label{lem:generically_free_subsingleton}
  \lean{AlgebraicGeometry.GenericFreeness.genericallyFree_of_subsingleton}
  \uses{def:generically_free}
  Let \(A\) be a non-trivial commutative ring.  The zero \(A\)-module is
  generically free over \(A\), with witness \(f = 1\).
\end{lemma}

\begin{proof}
\leanok
  The localisation of the zero module is zero, which is free (on the empty
  basis), and \(1 \ne 0\) since \(A\) is non-trivial.
\end{proof}

\begin{lemma}
\leanok
[Torsion base case]
  \label{lem:generically_free_torsion}
  \lean{AlgebraicGeometry.GenericFreeness.genericallyFree_of_torsion}
  \uses{def:generically_free}
  \source{nitsure-hilbert-quot}
  Let \(A\) be a domain and \(M\) a finite \(A\)-module every element of which
  is annihilated by some non-zero element of \(A\).  Then \(M\) is generically
  free over \(A\): indeed some localisation \(M_f\) with \(f \ne 0\) vanishes.
\end{lemma}

\begin{proof}
\leanok
  Choose a finite generating set \(m_1, \dots, m_r\) of \(M\) and non-zero
  annihilators \(a_i m_i = 0\).  The product \(f := a_1 \cdots a_r\) is
  non-zero (\(A\) is a domain) and annihilates every generator (extract the
  factor \(a_i\) and commute), hence — the annihilator of \(f\) being a
  submodule containing the generators — annihilates all of \(M\).  Every
  element of \(M_f\) is then zero, and the zero module is free.
\end{proof}

\begin{lemma}
\leanok
[Generic freeness: finite modules]
  \label{lem:generically_free_finite}
  \lean{AlgebraicGeometry.GenericFreeness.exists_free_localizationAway_of_finite}
  \uses{def:generically_free}
  Let \(A\) be a noetherian domain and \(M\) a finite \(A\)-module.  Then \(M\)
  is generically free over \(A\).
\end{lemma}

\begin{proof}
\leanok
  A finite module over a noetherian ring is finitely presented, and
  \(M \otimes_A \mathrm{Frac}(A)\) is a finite-dimensional
  \(\mathrm{Frac}(A)\)-vector space, hence free.  Freeness at the generic
  point of a finitely presented module spreads out to a basic open: there is
  \(f \ne 0\) in \(A\) with \(M_f\) free over \(A_f\).
\end{proof}

\begin{lemma}
\leanok
[Generic flatness: finite modules]
  \label{lem:generically_flat_finite}
  \lean{AlgebraicGeometry.GenericFreeness.exists_flat_localizationAway_of_finite}
  \uses{def:generically_free}
  Let \(A\) be a noetherian domain and \(M\) a finite \(A\)-module.  Then there
  is \(f \ne 0\) in \(A\) with \(M_f\) flat over \(A_f\).
\end{lemma}

\begin{proof}
\leanok
  \uses{lem:generically_free_finite}
  Immediate from \cref{lem:generically_free_finite}: free modules are flat.
\end{proof}

\begin{lemma}
\leanok
[Generic freeness: the module-finite case]
  \label{lem:generically_free_module_finite}
  \lean{AlgebraicGeometry.GenericFreeness.exists_free_localizationAway_of_moduleFinite}
  \uses{def:generically_free}
  Let \(A\) be a noetherian domain, \(B\) a module-finite \(A\)-algebra, and
  \(M\) a finite \(B\)-module.  Then \(M\) is generically free over \(A\).
  (This is the case of \cref{thm:generic_flatness_algebraic} where the
  morphism \(\Spec B \to \Spec A\) is finite.)
\end{lemma}

\begin{proof}
\leanok
  \uses{lem:generically_free_finite}
  A finite module over a module-finite \(A\)-algebra is a finite
  \(A\)-module, so \cref{lem:generically_free_finite} applies.
\end{proof}

\begin{lemma}
\leanok
[Generic-rank comparison for a finite module over a domain]
  \label{lem:generic_rank_comparison}
  \lean{AlgebraicGeometry.GenericFreeness.exists_generic_rank_comparison}
  Let \(D\) be a domain and \(M\) a finite \(D\)-module.  Then there exist
  \(r \in \mathbb{N}\), an injective \(D\)-linear map
  \(\Phi : D^{\oplus r} \to M\),
  and a non-zero \(d \in D\) annihilating the cokernel
  \(M / \Phi(D^{\oplus r})\); that is, \(M\) sits in an exact sequence
  \(0 \to D^{\oplus r} \to M \to T \to 0\) with \(d\,T = 0\).
\end{lemma}

\begin{proof}
\leanok
  Localise at \(S := D \setminus \{0\}\): the localisation \(S^{-1}M\) is a
  finite-dimensional vector space over \(\mathrm{Frac}(D) = S^{-1}D\); let
  \(r\) be its dimension.  Every element of \(S^{-1}M\) has the form
  \(m/s\), so scaling each vector of a basis by the (unit) image of its
  denominator produces a basis \((\bar c_1, \dots, \bar c_r)\) whose members
  are images of elements \(c_1, \dots, c_r \in M\); let
  \(\Phi(e_i) := c_i\).  A \(D\)-linear relation among the \(c_i\)
  localises to a \(\mathrm{Frac}(D)\)-linear relation among the basis
  vectors \(\bar c_i\), whose coefficients — images of the original ones —
  must vanish, and \(D \to \mathrm{Frac}(D)\) is injective, so \(\Phi\) is
  injective.  For the cokernel: given \(y \in M\), expand its image in the
  basis with coefficients in \(\mathrm{Frac}(D)\) and clear their
  denominators by a single \(s \in S\); then \(s \cdot y - \Phi(u)\) is in
  the kernel of \(M \to S^{-1}M\) for some \(u \in D^{\oplus r}\), hence is
  killed by some \(t \in S\), so \(ts \ne 0\) annihilates the class of
  \(y\) in the cokernel \(T\).  Finally \(T\) is a finite \(D\)-module, so
  the product of such annihilators over a finite generating set is a single
  non-zero uniform annihilator \(d\).
\end{proof}

\begin{lemma}
\leanok
[Normalisation core: generic freeness of quotient domains]
  \label{lem:generically_free_domain_core}
  \lean{AlgebraicGeometry.GenericFreeness.genericallyFree_quotient_prime}
  \uses{def:generically_free}
  \source{nitsure-hilbert-quot}
  Let \(A\) be a noetherian domain, \(B\) a finite-type \(A\)-algebra, and
  \(\pp \subset B\) a prime ideal.  Then the quotient domain \(B/\pp\) is
  generically free as an \(A\)-module.
\end{lemma}

\begin{proof}
\leanok
  \uses{lem:generically_free_torsion, lem:generically_free_transport,
    lem:generically_free_splice, lem:generic_rank_comparison}
  Write \(C := B/\pp\), a finite-type \(A\)-algebra which is a domain, and let
  \(K := \mathrm{Frac}(A)\).  If the kernel of \(A \to C\) is non-zero, every
  element of \(C\) is annihilated by any non-zero \(a \in \ker(A \to C)\)
  (as \(a \cdot c = (a \cdot 1)\, c = 0\)), and
  \cref{lem:generically_free_torsion} concludes.  So assume \(A \subseteq C\).
  Then \(C_K := K \otimes_A C\) is a non-zero finite-type \(K\)-algebra domain;
  let \(m\) be its Krull dimension.  We induct on \(m\), the statement for all
  smaller values (over all noetherian domains \(A\)) being granted.

  By Noether normalisation there are \(b_1, \dots, b_m \in C_K\), algebraically
  independent over \(K\), with \(C_K\) module-finite over
  \(K[b_1, \dots, b_m]\); scaling by non-zero elements of \(A\) (which does not
  affect either property) we may take \(b_i \in C\).  Fix a finite generating
  set of \(C\) as an \(A\)-algebra; each generator satisfies a monic equation
  over \(K[b_1, \dots, b_m]\), and clearing the finitely many denominators of
  all coefficients produces a single \(g \ne 0\) in \(A\) such that each
  generator of \(C_g\) is integral over \(A_g[b_1, \dots, b_m]\) (multiply the
  monic equation for \(x\) through by \(g^{\deg}\) and rewrite it as a monic
  equation for a unit multiple of \(x\)).  A finite-type algebra which is
  integral is module-finite, so \(C_g\) is a finite module over the polynomial
  ring \(D := A_g[b_1, \dots, b_m]\).

  Let \(r\) be the dimension of \(C_g \otimes_D \mathrm{Frac}(D)\) as a
  \(\mathrm{Frac}(D)\)-vector space and choose \(c_1, \dots, c_r \in C_g\)
  mapping to a basis.  The resulting map \(D^{\oplus r} \to C_g\) is injective
  (a \(D\)-linear relation among the \(c_i\) becomes a
  \(\mathrm{Frac}(D)\)-linear relation), and its cokernel \(T\) is a finite
  \(D\)-module with \(T \otimes_D \mathrm{Frac}(D) = 0\), i.e.\ torsion;
  this generic-rank packaging, including the uniform annihilator below, is
  \cref{lem:generic_rank_comparison}.  A
  non-zero \(d \in D\) annihilating \(T\) exhibits
  \(T_K := K \otimes_{A_g} T\) as a finite module over
  \(K[b_1, \dots, b_m]/(d')\) for the non-zero image \(d'\) of \(d\), a ring of
  Krull dimension \(< m\); presenting that ring as a quotient of a polynomial
  ring and filtering as in \cref{thm:generic_flatness_algebraic}'s dévissage,
  the inductive hypothesis (over the noetherian domain \(A_g\)) together with
  \cref{lem:generically_free_splice} yields \(h \ne 0\) in \(A_g\), which we
  may take in \(A\), with \(T_h\) free over \((A_g)_h\).  Localising the exact
  sequence \(0 \to D^{\oplus r} \to C_g \to T \to 0\) at \(h\) and splitting
  off the free (hence projective) quotient \(T_h\), we get
  \[
    C_{gh} \;\cong\; \bigl(A_{gh}[b_1, \dots, b_m]\bigr)^{\oplus r} \oplus T_h ,
  \]
  a direct sum of free \(A_{gh}\)-modules (a polynomial ring is free on its
  monomials), so \(f := gh\) is a generic-freeness witness for \(C\)
  (\cref{lem:generically_free_transport} normalising the witnesses).
\end{proof}

\begin{theorem}
\leanok
[Generic flatness, algebraic form]
  \label{thm:generic_flatness_algebraic}
  \lean{AlgebraicGeometry.GenericFreeness.genericFlatnessAlgebraic}
  \uses{def:generically_free}
  \source{nitsure-hilbert-quot}
  \textit{Source: [Nitsure], \S4, ``Lemma on Generic Flatness''.}
  Let \(A\) be a noetherian domain, \(B\) a finite-type \(A\)-algebra, and \(M\) a finite
  \(B\)-module. Then \(M\) is generically free over \(A\): there exists
  \(f \in A\), \(f \ne 0\), such that the localisation
  \(M_f\) is a free \(A_f\)-module.
\end{theorem}

\begin{proof}
\leanok
  \uses{lem:generically_free_subsingleton, lem:generically_free_of_equiv,
    lem:generically_free_splice, lem:generically_free_domain_core}
  The ring \(B\) is noetherian (it is of finite type over the noetherian
  \(A\)), so the finite \(B\)-module \(M\) admits a finite prime filtration
  \[
    0 = M_0 \subset M_1 \subset \cdots \subset M_r = M
  \]
  with successive quotients \(M_i / M_{i-1} \cong B / \mathfrak{p}_i\) for
  prime ideals \(\mathfrak{p}_i \subset B\).  By induction on the filtration
  length it suffices that generic freeness over \(A\) holds for the zero
  module (\cref{lem:generically_free_subsingleton}), holds for each quotient
  domain \(B/\mathfrak{p}\) (\cref{lem:generically_free_domain_core},
  transported along the isomorphism \(M_i/M_{i-1} \cong B/\mathfrak{p}_i\) by
  \cref{lem:generically_free_of_equiv}), and is stable under extensions
  (the splicing lemma, \cref{lem:generically_free_splice}, applied to
  \(0 \to M_{i-1} \to M_i \to M_i/M_{i-1} \to 0\)).
\end{proof}

\section{Geometric glue: from generic freeness to generic flatness}

The algebraic statement concerns one module over one algebra; the geometric
statement quantifies over all affine pairs below the witness open. The glue
consists of a mixed-base stability lemma for flatness under module
localization, two restriction-compatibility lemmas for the induced module
structures, and a two-layer basic-open reduction.

\begin{lemma}[Mixed-base stability of flatness under module localization]
  \leanok
  \label{lem:flat_localizedModule_algebra}
  \lean{Module.Flat.of_isLocalizedModule_algebra}
  Let \(B\) be an algebra over a commutative ring \(R\), let \(N\) be a
  \(B\)-module which is flat as an \(R\)-module, and let
  \(g : N \to N'\) be a localization of \(B\)-modules at a submonoid
  \(p \subseteq B\). Then \(N'\) is flat as an \(R\)-module.
\end{lemma}
\begin{proof}
  \leanok
  Let \(f : Q \hookrightarrow Q'\) be an injective map of \(R\)-modules. The
  square
  \[
    \begin{array}{ccc}
      N \otimes_R Q & \longrightarrow & N \otimes_R Q' \\
      \downarrow & & \downarrow \\
      N' \otimes_R Q & \longrightarrow & N' \otimes_R Q'
    \end{array}
  \]
  identifies the bottom map \(\mathrm{id}_{N'} \otimes f\) with the
  localization at \(p\) of the top map \(\mathrm{id}_N \otimes f\): indeed
  \(N' \otimes_R Q\) is the localization of the \(B\)-module
  \(N \otimes_R Q\) at \(p\) (localization commutes with \(\, \otimes_R Q\),
  because \(S^{-1}(N \otimes_R Q) = S^{-1}B \otimes_B (N \otimes_R Q)
  = (S^{-1}B \otimes_B N) \otimes_R Q\)), and localizing the top map yields
  the bottom one. The top map is injective by flatness of \(N\) over \(R\),
  and localization of modules preserves injectivity; hence
  \(\mathrm{id}_{N'} \otimes f\) is injective.
\end{proof}

\begin{lemma}[Fibre-side restriction compatibility of \(\mathrm{appLE}\)]
  \leanok
  \label{lem:appLE_res_apply}
  \lean{AlgebraicGeometry.appLE_res_apply}
  Let \(p : X \to S\) be a morphism of schemes, \(U \subseteq S\) open,
  \(W' \subseteq W \subseteq p^{-1}(U)\) opens of \(X\), and
  \(u \in \Gamma(S, U)\). Then restricting
  \(p^\sharp_{U,W}(u) \in \Gamma(X, W)\) to \(W'\) yields
  \(p^\sharp_{U,W'}(u)\).
\end{lemma}
\begin{proof}
  \leanok
  Both sides are the composite of \(p^\sharp\) with restriction along
  \(W' \subseteq W \subseteq p^{-1}(U)\); this is the naturality of
  \(p^\sharp\) in the target open.
\end{proof}

\begin{lemma}[Base-side restriction compatibility of \(\mathrm{appLE}\)]
  \leanok
  \label{lem:appLE_base_res_apply}
  \lean{AlgebraicGeometry.appLE_base_res_apply}
  Let \(p : X \to S\) be a morphism of schemes, \(U' \subseteq U \subseteq S\)
  opens, \(W \subseteq p^{-1}(U')\) an open of \(X\), and
  \(u \in \Gamma(S, U)\). Then
  \(p^\sharp_{U',W}\bigl(u|_{U'}\bigr) = p^\sharp_{U,W}(u)\).
\end{lemma}
\begin{proof}
  \leanok
  Naturality of \(p^\sharp\) in the source open.
\end{proof}

\begin{lemma}[Section flatness transports along equal opens]
  \leanok
  \label{lem:flat_sections_congr}
  \lean{AlgebraicGeometry.flat_sections_congr}
  Let \(p : X \to S\), \(\mathcal F\) a module sheaf on \(X\),
  \(U_0 \subseteq S\) open, and \(O_1 = O_2 \subseteq p^{-1}(U_0)\) equal
  opens of \(X\). If \(\Gamma(\mathcal F, O_1)\) is flat over
  \(\Gamma(S, U_0)\) (via \(p^\sharp\)), then so is
  \(\Gamma(\mathcal F, O_2)\).
\end{lemma}
\begin{proof}
  \leanok
  Immediate: the two data coincide.
\end{proof}

\begin{lemma}[Chart-basic core of generic flatness]
  \leanok
  \label{lem:flat_section_chartBasic}
  \lean{AlgebraicGeometry.flat_section_chartBasic}
  \uses{def:coherent_sheaf_flat}
  Let \(p : X \to S\), let \(\mathcal F\) be quasi-coherent on \(X\), let
  \(U_0 \subseteq S\) be open, and let \(W_j \subseteq p^{-1}(U_0)\) be an
  affine chart. Let \(g \in A := \Gamma(S, U_0)\) be such that the localized
  section module \(\Gamma(\mathcal F, W_j)_g\) is free over \(A_g\). Then for
  every \(c \in \Gamma(X, W_j)\) with \(D(c) \subseteq p^{-1}(D(g))\), the
  module \(\Gamma(\mathcal F, D(c))\) is flat over \(A\).
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:flat_localizedModule_algebra, lem:appLE_res_apply,
    lem:flat_sections_congr, lem:qcoh_section_localization_basicOpen}
  Write \(\beta := p^\sharp_{U_0, W_j}(g) \in B_j := \Gamma(X, W_j)\), so that
  \(D(\beta) = W_j \cap p^{-1}(D(g))\). By the quasi-compact--quasi-separated
  section-localization theorem
  (\cref{lem:qcoh_section_localization_basicOpen}, applied to the affine
  \(W_j\)), the restriction
  \(\Gamma(\mathcal F, W_j) \to \Gamma(\mathcal F, D(\beta))\) is the
  localization of \(B_j\)-modules at the powers of \(\beta\); since
  \(\beta^k\) acts on these modules as \(g^k\) acts through \(p^\sharp\), it
  is also the localization of \(A\)-modules at the powers of \(g\). Hence
  \(\Gamma(\mathcal F, D(\beta)) \cong \Gamma(\mathcal F, W_j)_g\) is free,
  in particular flat, over \(A_g\), and therefore flat over \(A\) (flatness
  over \(A\) and over the localization \(A_g\) agree for
  \(A_g\)-modules).

  Since \(D(c) \subseteq D(\beta)\), the further restriction
  \(\Gamma(\mathcal F, D(\beta)) \to \Gamma(\mathcal F, D(c))\) is again a
  section localization (\cref{lem:qcoh_section_localization_basicOpen} on the
  affine \(D(\beta)\), at the restricted element \(c|_{D(\beta)}\), whose
  basic open is \(D(c)\)) — a localization of
  \(\Gamma(X, D(\beta))\)-modules. Mixed-base stability
  (\cref{lem:flat_localizedModule_algebra}, with base \(A\) and fibre algebra
  \(\Gamma(X, D(\beta))\)) shows \(\Gamma(\mathcal F, D(c|_{D(\beta)}))\) is
  flat over \(A\); the induced-structure compatibilities are
  \cref{lem:appLE_res_apply}, and \cref{lem:flat_sections_congr} transports
  the conclusion along \(D(c|_{D(\beta)}) = D(c)\).
\end{proof}

\begin{lemma}[Two-layer basic-open reduction]
  \leanok
  \label{lem:flat_section_pair}
  \lean{AlgebraicGeometry.flat_section_pair}
  \uses{def:coherent_sheaf_flat}
  Let \(p : X \to S\), \(\mathcal F\) quasi-coherent on \(X\),
  \(U_0 \subseteq S\) an affine open with \(A := \Gamma(S, U_0)\), and let
  \(\{W_c^{(j)}\}_{j}\) be affine charts contained in and covering
  \(p^{-1}(U_0)\). Let \(f \in A\) be such that each localized section
  module \(\Gamma(\mathcal F, W_c^{(j)})_f\) is free over \(A_f\). Then for
  every affine pair \(U \subseteq D(f)\), \(W \subseteq p^{-1}(U)\), the
  module \(\Gamma(\mathcal F, W)\) is flat over \(\Gamma(S, U)\) via
  \(p^\sharp\).
\end{lemma}
\begin{proof}
  \leanok
  \uses{lem:flat_section_chartBasic, lem:appLE_base_res_apply,
    lem:qcoh_section_localization_basicOpen, lem:generically_free_transport}
  For each \(x \in W\): choose (affine interchange on \(S\), applied to the
  affines \(U\) and \(U_0\) at the point \(p(x) \in U \cap U_0\)) sections
  \(a \in \Gamma(S, U)\), \(a' \in A\) with
  \(D(a) = D(a') \ni p(x)\); choose a chart \(W_c^{(j)} \ni x\); shrink to a
  basic open of \(W\) inside \(W \cap p^{-1}(D(a)) \cap W_c^{(j)}\)
  containing \(x\); and apply affine interchange on \(X\) to that basic open
  and the affine \(W_c^{(j)}\) to obtain \(b \in \Gamma(X, W)\),
  \(c \in \Gamma(X, W_c^{(j)})\) with
  \(x \in D(b) = D(c) \subseteq p^{-1}(D(a))\).

  The basic opens \(D(b)\) obtained this way cover \(W\), so the chosen
  elements \(b\) generate the unit ideal of \(\Gamma(X, W)\). Since each
  restriction \(\Gamma(\mathcal F, W) \to \Gamma(\mathcal F, D(b))\) is a
  localization of \(\Gamma(X, W)\)-modules at the powers of \(b\)
  (\cref{lem:qcoh_section_localization_basicOpen}), flatness of
  \(\Gamma(\mathcal F, W)\) over \(\Gamma(S, U)\) may be checked on the
  pieces \(\Gamma(\mathcal F, D(b))\) (flatness over the base is local with
  respect to a spanning family of the fibre algebra).

  Fix a piece, with its data \(a, a', j, c\). Freeness of
  \(\Gamma(\mathcal F, W_c^{(j)})_f\) localizes further to
  \(f a'\) (\cref{lem:generically_free_transport}), and
  \(D(c) = D(b) \subseteq p^{-1}(D(a)) = p^{-1}(D(a'))
  \subseteq p^{-1}(D(f a'))\) since \(D(a) \subseteq U \subseteq D(f)\).
  The chart-basic core (\cref{lem:flat_section_chartBasic}, at the element
  \(f a'\) and the chart section \(c\)) gives flatness of
  \(\Gamma(\mathcal F, D(b))\) over \(A\). Finally exchange the base ring
  twice along the localizations \(A \to \Gamma(S, D(a')) = \Gamma(S, D(a))\)
  and \(\Gamma(S, U) \to \Gamma(S, D(a))\) (for a module over a localized
  ring, flatness over the ring and over its localization agree; the
  compatibility of the induced structures is
  \cref{lem:appLE_base_res_apply}), obtaining flatness of
  \(\Gamma(\mathcal F, D(b))\) over \(\Gamma(S, U)\).
\end{proof}

\begin{theorem}[Generic flatness, geometric form]
  \leanok
  \label{thm:generic_flatness}
  \lean{AlgebraicGeometry.genericFlatness}
  \uses{thm:generic_flatness_algebraic}
  % SOURCE: [Nitsure], \S4, Theorem ``generic flatness''
  % (read from references/nitsure-hilbert-quot-src/nitsure-hilbert-quot.tex, L1781-L1787).
  % SOURCE QUOTE:
  % "Let \(S\) be a noetherian and integral scheme.
  %  Let \(p: X\to S\) be a finite type morphism, and let \(\F\) be a
  %  coherent sheaf of \({\mathcal O}_X\)-modules. Then there exists a non-empty
  %  open subscheme \(U\subset S\) such that the restriction of \(\F\)
  %  to \(X_U = p^{-1}(U)\) is flat over \({\mathcal O}_U\)."
  \textit{Source: [Nitsure], \S4, Theorem ``generic flatness''.}
  Let \(S\) be an integral locally noetherian scheme, \(p : X \to S\) a
  finite-type morphism (that is, quasi-compact and locally of finite type —
  quasi-compactness is essential: it bounds the number of denominators to
  clear, and the statement fails for the locally-of-finite-type morphism
  \(\coprod_q \Spec \mathbb{F}_q \to \Spec \Z\)), and \(\F\) a coherent
  \(\OO_X\)-module. Then there exists a non-empty open subscheme
  \(U \subseteq S\) such that the restriction of \(\F\) to
  \(X_U = p^{-1}(U)\) is \(\OO_U\)-flat.
\end{theorem}

% SOURCE QUOTE PROOF: "The above theorem has the following consequence, which
% follows by restricting attention to a non-empty affine open subscheme of \(S\)."
% (Source: nitsure-hilbert-quot-src/nitsure-hilbert-quot.tex, L1776-L1777. The
% Nitsure proof is the single sentence quoted: deduce the geometric form from
% the algebraic generic flatness statement (\cref{thm:generic_flatness_algebraic})
% by restricting to a non-empty
% affine open of \(S\), where the situation reduces to the algebraic statement.)
\begin{proof}
  \leanok
  \uses{thm:generic_flatness_algebraic, lem:flat_section_pair,
    lem:affine_finite_sections_nhds}
  Choose a non-empty affine open \(U_0 \subseteq S\); its section ring
  \(A := \Gamma(S, U_0)\) is a noetherian domain because \(S\) is integral and
  locally noetherian. The preimage \(p^{-1}(U_0)\) is quasi-compact
  (\(p\) is quasi-compact and \(U_0\) is), and every point of it has an
  affine open neighbourhood inside \(p^{-1}(U_0)\) on which the sections of
  \(\F\) are a finite module over the sections of \(\OO_X\)
  (\cref{lem:affine_finite_sections_nhds}); extract a finite subcover by
  affine charts \(W^{(1)}, \dots, W^{(r)}\). Each
  \(B_i := \Gamma(X, W^{(i)})\) is of finite type over \(A\) (as \(p\) is
  locally of finite type and both opens are affine) and each
  \(M_i := \Gamma(\F, W^{(i)})\) is a finite \(B_i\)-module, so algebraic
  generic freeness (\cref{thm:generic_flatness_algebraic}) produces
  \(f_i \in A \setminus \{0\}\) with \((M_i)_{f_i}\) free over \(A_{f_i}\).
  Set \(f := f_1 \cdots f_r \ne 0\); freeness localizes from each \(f_i\) to
  \(f\), so every \((M_i)_f\) is free over \(A_f\). The witness open is
  \(V := D(f) \subseteq U_0\), non-empty because \(S\) is reduced and
  \(f \ne 0\). The required flatness on every affine pair
  \(U \subseteq V\), \(W \subseteq p^{-1}(U)\) is exactly the two-layer
  basic-open reduction, \cref{lem:flat_section_pair}.
\end{proof}

\section{Existence of the flattening stratification}
\label{sec:flatstrat_exists}

\begin{theorem}

\leanok
[Existence of the flattening stratification]
  \label{thm:flattening_stratification_exists}
  \lean{AlgebraicGeometry.flatteningStratification}
  \uses{def:coherent_sheaf_flat, lem:flat_locus_open, lem:nonflat_locus_proper,
    lem:noetherian_induction_strata, thm:generic_flatness}

  % SOURCE: [Nitsure], \S4, Theorem ``flattening stratification''
  % (read from references/nitsure-hilbert-quot-src/nitsure-hilbert-quot.tex, L1811-L1841).
  % Parallel pointer: [Stacks Project], tag 052H (verbatim text not yet
  % retrieved into a local references/ file; see ``Notes for Plan Agent'').
  % SOURCE QUOTE:
  % "Let \(S\) be a noetherian scheme, and let
  %  \(\F\) be a coherent sheaf on the projective space \(\PP^n_S\)
  %  over \(S\). Then the set \(I\) of Hilbert polynomials of
  %  restrictions of \(\F\) to fibers of \(\PP^n_S\to S\)
  %  is a finite set. Moreover, for each \(f\in I\) there exist
  %  a locally closed subscheme \(S_f\) of \(S\),
  %  such that the following conditions are satisfied.
  %
  %  \textbf{(i) Point-set: } The underlying set \(|S_f|\) of \(S_f\) consists of
  %  all points \(s\in S\) where the Hilbert polynomial of the
  %  restriction of \(\F\) to \(\PP^n_s\) is \(f\). In particular,
  %  the subsets \(|S_f|\subset |S|\) are disjoint,
  %  and their set-theoretic union is \(|S|\).
  %
  %  \textbf{(ii) Universal property: }
  %  Let \(S'=\coprod S_f\) be the coproduct of the \(S_f\), and let
  %  \(i:S'\to S\) be the morphism induced by the inclusions
  %  \(S_f\hra S\). Then the sheaf \(i^*(\F)\) on \(\PP^n_{S'}\)
  %  is flat over \(S'\). Moreover, \(i:S'\to S\) has the universal property that
  %  for any morphism \(\varphi :T\to S\) the pullback \(\varphi^*(\F)\) on
  %  \(\PP^n_T\) is flat over \(T\) if and only if \(\varphi\) factors through
  %  \(i:S'\to S\).  The subscheme \(S_f\) is
  %  uniquely determined by the polynomial \(f\).
  %
  %  \textbf{(iii) Closure of strata: }
  %  Let the set \(I\) of Hilbert polynomials be given a total
  %  ordering, defined by putting \(f<g\) whenever \(f(n) < g(n)\) for all
  %  \(n \gg  0\). Then the closure in \(S\) of the subset \(|S_f|\)
  %  is contained in the union of all \(|S_g|\) where \(f\le g\)."
  \textit{Source: [Nitsure], \S4, Theorem ``flattening stratification'';
  cf.\ [Stacks Project], tag 052H.}
  Let \(S\) be a noetherian scheme and \(\F\) a coherent sheaf on the relative
  projective space \(\PP^n_S\) over \(S\).  Then:
  \begin{enumerate}[label=(\roman*)]
    \item The set \(I\) of Hilbert polynomials
      \(P_s(m) = \chi(\PP^n_s, \F_s(m))\) that arise as \(s\) ranges over \(S\) is
      finite.
    \item For each \(f \in I\) there exists a locally-closed subscheme
      \(S_f \hookrightarrow S\) whose underlying set \(|S_f|\) consists of exactly
      those \(s \in S\) with \(P_s = f\).  The collection \(\{|S_f|\}_{f \in I}\) is
      pairwise disjoint with set-theoretic union \(|S|\).  Forming the coproduct
      \(S' := \coprod_{f \in I} S_f\) and the morphism \(i : S' \to S\) assembled
      from the inclusions \(S_f \hookrightarrow S\), the pullback \(i^* \F\) on
      \(\PP^n_{S'}\) is flat over \(S'\), and \(i : S' \to S\) is universal with this
      property: for any morphism \(\varphi : T \to S\) such that \(\varphi^* \F\) on
      \(\PP^n_T\) is flat over \(T\), \(\varphi\) factors uniquely through \(i\).
      Each \(S_f\) is uniquely determined by \(f\).
    \item Order \(I\) by \(f < g \iff f(n) < g(n)\) for all \(n \gg 0\). Then the
      closure of \(|S_f|\) in \(|S|\) is contained in \(\bigcup_{g \ge f} |S_g|\).
  \end{enumerate}
\end{theorem}

% SOURCE QUOTE PROOF: "It is enough to prove the theorem for open subschemes of
% $S$ which cover $S$, as the resulting strata will then glue together by their
% universal property. [...] \textbf{General case: } We now allow the integer
% $n$ to be arbitrary. The idea of the proof is as follows: We show the
% existence of a stratification of $S$ which is a `g.c.d.' of the flattening
% stratifications for direct images $\pi_*\F(i)$ for all $i \ge N$ for some
% integer $N$ (where the flattening stratifications for $\pi_*\F(i)$ exist by
% case $n=0$ which we have treated above). This is the desired flattening
% stratification of $\F$ over $S$, as follows from Lemma \ref{graded module is flat}."
% (Source: nitsure-hilbert-quot-src/nitsure-hilbert-quot.tex, L1844-L1900; the
% full body L1844-L2090 is distributed across the sub-lemmas below, each of
% which carries its own verbatim % SOURCE QUOTE PROOF.)
\begin{proof}
  \uses{def:coherent_sheaf_flat, lem:flat_locus_open, lem:nonflat_locus_proper,
    lem:noetherian_induction_strata, thm:generic_flatness}
  Following [Nitsure], \S4, the construction reduces to the special case
  \(n = 0\), assembled over the relative twists \(\pi_* \F(N+i)\).  As the
  argument is local on \(S\) (the strata glue by their universal property, part
  (ii)), and as the three conclusions are proved through the sub-lemmas below,
  we assemble them in the order (i), (ii), (iii).

  \emph{Part (i): finiteness of the Hilbert-polynomial set.}
  By Lemma~\ref{lem:nonflat_locus_proper}, Noetherian induction on \(S\) (whose
  inductive step invokes generic flatness, Theorem~\ref{thm:generic_flatness},
  on the integral locally-closed strata cut out by the irreducible-component
  decomposition of \(S\)) produces finitely many reduced, pairwise disjoint,
  locally-closed subschemes \(V_i \hookrightarrow S\) set-theoretically covering
  \(|S|\) on each of which \(\F|_{\PP^n_{V_i}}\) is flat over \(\OO_{V_i}\) in
  the sense of Definition~\ref{def:coherent_sheaf_flat}.  On a flat family the
  Hilbert polynomial \(P_s(m) = \chi(\PP^n_s, \F_s(m))\) is locally constant, so
  it takes finitely many values on each quasi-compact \(V_i\); the finite union
  over the finitely many \(V_i\) shows that the set \(I\) of Hilbert polynomials
  occurring on \(S\) is finite.

  \emph{Part (ii): existence of the strata and their universal property.}
  Working over the flat pieces \(V_i\) of part (i), semicontinuity
  (cohomology-and-base-change) yields a uniform integer \(N\) such that, for all
  \(s \in S\) and \(m \ge N\), the higher cohomologies \(H^r(\PP^n_s, \F_s(m))\)
  vanish for \(r \ge 1\) and the base-change homomorphism
  \((\pi_* \F(m))|_s \to H^0(\PP^n_s, \F_s(m))\) is an isomorphism.  With this
  \(N\) fixed, Lemma~\ref{lem:noetherian_induction_strata} applies the \(n = 0\)
  stratification mechanism (Lemma~\ref{lem:flat_locus_open}) successively to the
  coherent sheaves \(E_i = \pi_* \F(N+i)\) on \(S\) for \(i = 0, \dots, n\),
  refining stratum by stratum.  After \(n+1\) steps this produces a finite
  family of locally-closed subschemes \(W_{e_0, \dots, e_n} \subseteq S\),
  universal for the property that each \(E_i\) pulls back to a locally-free sheaf
  of constant rank \(e_i\); re-indexing the tuple \((e_0, \dots, e_n)\) by the
  unique numerical polynomial \(f\) of degree \(\le n\) with \(f(N+i) = e_i\)
  gives the family \(\{W_f\}_{f \in I}\).  At any \(s \in W_f\) the base-change
  isomorphism and the cohomology vanishing force \(P_s = f\) (since \(P_s\) has
  degree \(\le n\) and is determined by its values at \(N, \dots, N+n\)), so
  \(|W_f| = \{s : P_s = f\}\); these subsets are disjoint with union \(|S|\).
  The required locally-closed subscheme \(S_f \subseteq W_f\) is the closed
  subscheme cut out by the further local-freeness conditions
  ``\(\pi_* \F(N+i)|_{W_f}\) is locally free of rank \(f(N+i)\) for all
  \(i \ge 0\)'' (a chain of coherent ideal sheaves that stabilises by
  Noetherianity), and \(\F|_{\PP^n_{S_f}}\) is flat over \(S_f\)
  (Definition~\ref{def:coherent_sheaf_flat}) because the graded module
  \(\bigoplus_{i \gg 0} \pi_* \F(i)|_{S_f}\) is locally free.  Forming
  \(S' = \coprod_{f \in I} S_f\) and the assembled morphism \(i : S' \to S\),
  the pullback \(i^* \F\) is \(S'\)-flat; and any \(\varphi : T \to S\) with
  \(\varphi^* \F\) flat over \(T\) has locally constant Hilbert polynomial, hence
  decomposes \(T = \coprod_f T_f\) into the open-closed loci where the polynomial
  is \(f\), each factoring uniquely through the corresponding \(S_f\) by the
  local-freeness universal property of Lemma~\ref{lem:flat_locus_open}.  This
  gives the unique factorisation through \(i\), and shows each \(S_f\) is
  uniquely determined by \(f\).

  \emph{Part (iii): closure of strata.}
  Order \(I\) by \(f < g \iff f(n) < g(n)\) for all \(n \gg 0\).  Since each
  \(S_f\) is a stratum of the \(n = 0\) flattening stratification applied to a
  single direct image \(\pi_* \F(p)\) for \(p \gg 0\) separating all polynomials
  in \(I\), the closure-of-strata assertion reduces to the upper-semicontinuity
  statement already established in the \(n = 0\) case
  (Lemma~\ref{lem:flat_locus_open}): the closure of \(|S_f|\) in \(|S|\) is
  contained in \(\bigcup_{g \ge f} |S_g|\).
\end{proof}

The proof is long; we split it into sub-lemmas below.  The first sub-lemma is
the special case \(n = 0\) (when \(\PP^n_S = S\) and \(\F\) is just a coherent sheaf
on \(S\)), which already contains the flat-locus-as-locally-closed-subscheme
mechanism. The remaining inputs are (A) the Noetherian-induction reduction to
the flat case via generic flatness; (B) uniform vanishing of higher cohomology
of \(\F(m)\) for \(m \gg 0\) on \(\F_{V_i}\) for the flat pieces; and (C) base-change
isomorphisms for the direct images \(\pi_* \F(m)\), which allow comparing
strata for different twists.

\begin{lemma}
[Flat locus on \(S\) for a coherent sheaf is a locally-closed stratification]
  \label{lem:flat_locus_open}
  \lean{AlgebraicGeometry.TODO.flatLocusOpen}

  % SOURCE: [Nitsure], \S4, Theorem ``flattening stratification'', Special case
  % \(n=0\) (read from references/nitsure-hilbert-quot-src/nitsure-hilbert-quot.tex,
  % L1849-L1885).
  % SOURCE QUOTE:
  % "\textbf{Special case: } Let \(n=0\), so that \(\PP^n_S=S\).
  %  For any \(s\in S\), the \textbf{fiber} \(\F|_s\) of \(\F\) over \(s\)
  %  will mean the pull-back of \(\F\) to the subscheme
  %  \(\spec \kappa(s)\) [...]. The Hilbert polynomial of the
  %  restriction of \(\F\) to the fiber over \(s\) is the degree \(0\)
  %  polynomial \(e \in \Q[\lambda]\), where \(e = \dim_{\kappa(s)}\F|_s\).
  %
  %  By Nakayama lemma, any basis of \(\F|_s\) prolongs to a neighbourhood \(U\) of \(s\)
  %  to give a set of generators for \(F|_U\). Repeating this argument,
  %  we see that there exists a smaller neighbourhood \(V\) of \(s\) in which
  %  there is a right-exact sequence
  %  ${\mathcal O}_V^{\oplus m} \stackrel{\psi}{\to} {\mathcal O}_V^{\oplus e}
  %  \stackrel{\phi}{\to} \F \to 0$.
  %  Let \(I_e\subset {\mathcal O}_V\) be the ideal sheaf
  %  formed by the entries of the \(e\times m\) matrix
  %  \((\psi_{i,j})\) [...].
  %  Let \(V_e\) be the closed subscheme of \(V\) defined by \(I_e\).
  %  For any morphism of schemes
  %  \(f: T\to V\), the pull-back sequence [...]
  %  is exact, by right-exactness of tensor products.
  %  Hence the pull-back \(f^*\F\) is a locally free \({\mathcal O}_T\)-module of
  %  rank \(e\) if and only if \(f^*\psi =0\), that is,
  %  \(f\) factors via the subscheme \(V_e\hra V\) defined by the vanishing of all
  %  entries \(\psi_{i,j}\).
  %  Thus we have proved assertions \textbf{(i)} and \textbf{(ii)} of the theorem.
  %
  %  As the rank of the matrix \((\psi_{i,j})\) is lower semi-continuous,
  %  it follows that the function \(e\) is upper semi-continuous, which
  %  proves the assertion \textbf{(iii)} of the theorem, completing its proof
  %  when \(n=0\)."
  \textit{Source: [Nitsure], \S4 (special case \(n=0\) of the main theorem).}
  Let \(S\) be a noetherian scheme and let \(\F\) be a coherent \(\OO_S\)-module.
  For each integer \(e \ge 0\), there exists a locally-closed subscheme
  \(S_e \hookrightarrow S\) whose underlying set is
  \(\{s \in S : \dim_{\kappa(s)} \F|_s = e\}\). The family \(\{S_e\}_e\) is a
  finite, mutually disjoint, set-theoretically covering stratification of \(S\),
  and a morphism \(f : T \to S\) pulls back \(\F\) to a locally-free
  \(\OO_T\)-module of rank \(e\) if and only if \(f\) factors through \(S_e\).  The
  function \(s \mapsto \dim_{\kappa(s)} \F|_s\) is upper-semicontinuous on \(S\).
\end{lemma}

\begin{proof}
  
  By [Nitsure], \S4 (verbatim above): for \(s \in S\), Nakayama produces a
  neighbourhood \(V\) of \(s\) and a right-exact local presentation
  \(\OO_V^{\oplus m} \xrightarrow{\psi} \OO_V^{\oplus e} \to \F|_V \to 0\) where
  \(e = \dim_{\kappa(s)} \F|_s\).  The closed subscheme \(V_e \hookrightarrow V\)
  defined by the ideal sheaf generated by the matrix entries of \(\psi\)
  represents the locus where \(\F\) becomes locally free of rank \(e\): by
  right-exactness of tensor product, a base change \(f : T \to V\) has
  \(f^* \F\) locally free of rank \(e\) iff \(f^* \psi = 0\), iff \(f\) factors
  through \(V_e\).  Glueing the \(V_e\) over the cover of \(S\) by such
  neighbourhoods (and intersecting with \(V \setminus \bigcup_{e' > e} V_{e'}\)
  to get the strictly-equal-to-\(e\) stratum) produces the locally-closed
  \(S_e\). Upper-semicontinuity of \(e\) follows because the rank of \(\psi\) is
  lower-semicontinuous.
\end{proof}

\begin{lemma}
[Noetherian induction reduction to the flat case]
  \label{lem:nonflat_locus_proper}
  \lean{AlgebraicGeometry.TODO.nonflatLocusProper}
  \uses{thm:generic_flatness}

  % SOURCE: [Nitsure], \S4, Theorem ``flattening stratification'', General case
  % opening paragraphs (read from
  % references/nitsure-hilbert-quot-src/nitsure-hilbert-quot.tex, L1888-L1929).
  % SOURCE QUOTE:
  % "\textbf{General case: } We now allow the integer \(n\) to be arbitrary.
  %  The idea of the proof is as follows:
  %  We show the existence of a stratification
  %  of \(S\) which is a `g.c.d.' of the flattening
  %  stratifications for direct images \(\pi_*\F(i)\) for all \(i \ge N\) for some
  %  integer \(N\) [...].
  %
  %  As \(S\) is noetherian, it is a finite union of irreducible components,
  %  and these are closed in \(S\).
  %  Let \(Y\) be an irreducible component of \(S\), and let \(U\) be the
  %  non-empty open subset of \(Y\) which consists of all points which do not
  %  lie on any other irreducible component of \(S\).
  %  Let \(U\) be given the reduced subscheme
  %  structure. Note that this makes \(U\) an integral scheme, which is a
  %  locally closed subscheme of \(S\).
  %  By Theorem \ref{generic flatness} on generic flatness,
  %  \(U\) has a non-empty open subscheme \(V\) such that
  %  the restriction of \(\F\) to \(\PP^n_V\) is flat over \({\mathcal O}_V\).
  %  Now repeating the argument with \(S\) replaced by its reduced
  %  closed subscheme \(S-V\),
  %  it follows by noetherian induction on \(S\) that there exist
  %  finitely many reduced, locally closed, mutually disjoint subschemes
  %  \(V_i\) of \(S\) such that set-theoretically
  %  \(|S|\) is the union of the \(|V_i|\) and
  %  the restriction of \(\F\) to \(\PP^n_{V_i}\) is flat over \({\mathcal O}_{V_i}\).
  %  As each \(V_i\) is a noetherian scheme, and as the Hilbert polynomials
  %  are locally constant for a flat family of sheaves, it follows
  %  that only finitely many polynomials occur in \(V_i\)
  %  in the family of Hilbert polynomials \(P_s(m) = \chi(\PP^n_s,\F_s(m))\)
  %  as \(s\) varies over points of \(V_i\). This allows us to conclude
  %  the following:
  %
  %  \textbf{(A) } Only finitely many distinct Hilbert polynomials
  %  \(P_s(m) = \chi(\PP^n_s,\F_s(m))\) occur, as \(s\)
  %  varies over all of \(S\)."
  \textit{Source: [Nitsure], \S4 (general case, reduction by Noetherian induction).}
  Let \(S\) be a noetherian scheme, \(\pi : \PP^n_S \to S\) the relative projective
  space, and \(\F\) a coherent sheaf on \(\PP^n_S\).  Then there exist finitely many
  reduced, locally-closed, pairwise disjoint subschemes \(V_i \hookrightarrow S\)
  whose set-theoretic union is \(|S|\) and such that \(\F|_{\PP^n_{V_i}}\) is
  flat over \(\OO_{V_i}\) for each \(i\). Consequently, only finitely many
  Hilbert polynomials \(P_s(m) = \chi(\PP^n_s, \F_s(m))\) occur as \(s\) varies
  over \(S\).
\end{lemma}

\begin{proof}
  
  By [Nitsure], \S4: \(S\) being noetherian, decompose \(|S| = \bigcup Y_j\) into
  finitely many irreducible components, each closed.  For an irreducible
  component \(Y\), let \(U \subseteq Y\) be the open subset not lying on any other
  component, given the reduced subscheme structure (an integral, locally-closed
  subscheme of \(S\)).  By \cref{thm:generic_flatness} applied to
  \(\F|_{\PP^n_U}\), there is a non-empty open \(V \subseteq U\) over which \(\F\) is
  flat.  Replace \(S\) by the reduced closed complement \(S \setminus V\) and
  repeat. Noetherianity of \(S\) terminates the induction in finitely many
  steps, producing the \(V_i\).  On each \(V_i\) the Hilbert polynomial is locally
  constant (flat families have locally constant Hilbert polynomial), hence
  takes finitely many values on the noetherian (hence quasi-compact) scheme
  \(V_i\); the finite union of finite sets is finite.
\end{proof}

\begin{lemma}
[Assembly of the strata by Noetherian induction]
  \label{lem:noetherian_induction_strata}
  \lean{AlgebraicGeometry.TODO.noetherianInductionStrata}
  \uses{lem:flat_locus_open, lem:nonflat_locus_proper, thm:flat_base_change_cohomology}

  % SOURCE: [Nitsure], \S4, Theorem ``flattening stratification'', Assembly via
  % direct images \(E_i = \pi_*\F(N+i)\) (read from
  % references/nitsure-hilbert-quot-src/nitsure-hilbert-quot.tex, L1992-L2090).
  % SOURCE QUOTE:
  % "Let \(\pi : \PP^n_S \to S\) denote the projection. Consider the
  %  coherent sheaves \(E_0, \ldots, E_n\) on \(S\), defined by
  %  \(E_i = \pi_*\F(N+i)\mbox{ for } i=0,\ldots,n.\)
  %  By applying the special case of of the theorem (where the relative dimension
  %  \(n\) of \(\PP^n_S\) is \(0\)) to the sheaf \(E_0\) on \(\PP^0_S=S\), we get
  %  a stratification \((W_{e_0})\) of \(S\)
  %  indexed by integers \(e_0\), such that for any morphism
  %  \(f: T\to S\) the pull-back \(f^*E_0\) is a locally free \({\mathcal O}_T\)-module
  %  of rank \(e_0\) if and only if \(f\) factors via \(W_{e_0} \hra S\).
  %  Next, for each stratum \(W_{e_0}\), we take the
  %  flattening stratification \((W_{e_0,e_1})\) for \(E_1|_{W_{e_0}}\), and so on.
  %  Thus in \(n+1\) steps, we obtain finitely many
  %  locally closed subschemes
  %  \(W_{e_0,\ldots,e_n} \subset S\)
  %  such that for any morphism \(f:T\to S\) the pull-back
  %  \(f^*E_i\) for \(i=0,\ldots,n\) is a locally free \({\mathcal O}_T\)-modules of
  %  of constant rank \(e_i\) if and only if \(f\) factors
  %  via \(W_{e_0,\ldots,e_n} \hra S\). [...]
  %  This shows that at any point \(s\in W_f\), the Hilbert polynomial
  %  \(P_s(m)\) equals \(f\). The desired locally closed subscheme \(S_f \subset S\),
  %  whose existence is asserted by the theorem,
  %  will turn out to be a certain closed subscheme \(S_f \subset W_f\) whose
  %  underlying subset is all of \(|W_f|\). [...]
  %  As \(S_f\) is the flattening stratification for \(\pi_*\F(p)\),
  %  the property (iii) of the theorem follows from the corresponding property in
  %  the case \(n=0\), applied to the sheaf \(\pi_*\F(p)\) on \(S\)."
  \textit{Source: [Nitsure], \S4 (assembly step combining the \(n=0\) special case
  applied to direct images \(\pi_*\F(N+i)\)).}
  Let \(S\), \(\pi : \PP^n_S \to S\), \(\F\) be as in
  \cref{thm:flattening_stratification_exists}.  There exists an integer
  \(N \ge 0\) such that for every \(s \in S\) the higher cohomologies
  \(H^r(\PP^n_s, \F_s(m))\) vanish for all \(r \ge 1\), \(m \ge N\), and the base-change
  homomorphism \((\pi_* \F(m))|_s \to H^0(\PP^n_s, \F_s(m))\) is an isomorphism
  for \(m \ge N\), \(s \in S\).  Iterating the \(n=0\) stratification
  (\cref{lem:flat_locus_open}) on the coherent sheaves \(E_i := \pi_* \F(N+i)\)
  for \(i = 0, \dots, n\) produces a finite collection of locally-closed
  subschemes \(W_{e_0, \dots, e_n} \subseteq S\) indexed by tuples
  \((e_0, \dots, e_n) \in \Z^{n+1}\). Re-indexing each tuple by the unique
  numerical polynomial \(f \in \Q[\lambda]\) of degree \(\le n\) with
  \(f(N+i) = e_i\) gives a finite stratification \(\{W_f\}\) of \(S\), and the
  locally-closed subschemes \(S_f \subseteq W_f\) of
  \cref{thm:flattening_stratification_exists} are obtained by further
  intersecting with the closed-subscheme conditions that
  \(\pi_* \F(N+i)|_{W_f}\) be locally free of rank \(f(N+i)\) for all \(i \ge 0\)
  (terminating by Noetherianity).
\end{lemma}

% SOURCE QUOTE PROOF: full proof is the long body of [Nitsure], \S4, Theorem
% ``flattening stratification'', spanning L1844-L2090 (~250 lines: special
% case, Noetherian induction, statements (A)/(B)/(C), then the assembly via
% direct images and the closed-subscheme construction inside each \(W_f\)).
% Verbatim retrieved per-sub-lemma above; the structural restatement below in
% project notation is the canonical Nitsure-style restated proof.
\begin{proof}
  
  By \cref{lem:nonflat_locus_proper}, the family of Hilbert polynomials
  \(\{P_s\}\) on \(S\) is finite, and there is a finite stratification
  \(\{V_i\}\) of \(S\) on each of which \(\F\) is flat with locally-constant
  Hilbert polynomial.  Apply semi-continuity (cohomology and base change,
  [Hartshorne], III.12 / [Stacks Project], tag 02KH for the \(H^0\)-version) on
  the noetherian \(V_i\) to obtain a uniform integer \(N_1\) with
  \(R^r \pi_* \F(m) = 0\) for \(r \ge 1\), \(m \ge N_1\), and \(H^r(\PP^n_s, \F_s(m)) = 0\)
  for all \(s \in S\), \(r \ge 1\), \(m \ge N_1\) (statement (B) of Nitsure's proof).
  Choose \(r_i \ge N_1\) on each \(V_i\) such that the base-change comparison
  \((\pi_* \F(m))|_{V_i} \to (\pi_i)_* \F_{V_i}(m)\) is an isomorphism for
  \(m \ge r_i\); composing with the cohomology-and-base-change isomorphism
  \(((\pi_i)_* \F_{V_i}(m))|_s \xrightarrow{\sim} H^0(\PP^n_s, \F_s(m))\) for
  the flat family on \(V_i\) yields uniform \(N = \max r_i\) such that
  \((\pi_* \F(m))|_s \to H^0(\PP^n_s, \F_s(m))\) is an isomorphism for \(m \ge N\)
  and all \(s \in S\) (statement (C)).

  Apply \cref{lem:flat_locus_open} sequentially to the coherent sheaves
  \(E_i = \pi_* \F(N+i)\) on \(S\) for \(i = 0, \dots, n\): \(E_0\) produces a
  stratification \(\{W_{e_0}\}\) of \(S\); on each \(W_{e_0}\) apply the lemma to
  \(E_1|_{W_{e_0}}\) to refine to \(\{W_{e_0, e_1}\}\), and continue. After \(n+1\)
  steps the \(W_{e_0, \dots, e_n}\) form a finite locally-closed stratification
  of \(S\) universal for the property ``each \(E_i\) pulls back to a locally-free
  \(\OO_T\)-module of constant rank \(e_i\)''.  Re-index \((e_0, \dots, e_n)\) by the
  unique numerical polynomial \(f\) of degree \(\le n\) with \(f(N+i) = e_i\) to get
  the family \(\{W_f\}\) indexed by polynomials.  At any \(s \in W_f\), statement
  (C) and vanishing (B) force \(P_s = f\) (since \(P_s\) has degree \(\le n\) and is
  determined by its values at \(N, N+1, \dots, N+n\)).

  The Hilbert-polynomial-coincidence locus \(|W_f|\) may have non-flat scheme
  structure; refine to \(S_f \subseteq W_f\) by intersecting with the
  closed-subscheme conditions ``\(\pi_* \F(N+i)|_{W_f}\) is locally-free of rank
  \(f(N+i)\)'' for all \(i \ge 0\).  These conditions are cut out by an increasing
  chain of coherent ideal sheaves \(I_0 \subseteq I_0 + I_1 \subseteq \cdots\) on
  \(W_f\); Noetherianity terminates the chain in finitely many steps, defining a
  coherent ideal sheaf \(I\) and a closed subscheme \(S_f \subseteq W_f\) with
  \(|S_f| = |W_f|\) on which \(\pi_* \F(N+i)\) is locally-free of rank \(f(N+i)\)
  for all \(i \ge 0\). The base-change-without-flatness lemma (cf.\ [Nitsure],
  \S3) yields some \(N' \ge N\) such that for \(i \ge N'\),
  \((\pi_* \F(i))|_{S_f} \to (\pi_{S_f})_* \F_{S_f}(i)\) is an isomorphism, hence
  the graded \(A\)-module \(\bigoplus_{i \ge N'} (\pi_{S_f})_* \F_{S_f}(i)\) is
  locally-free on \(S_f\), hence \(\F_{S_f}\) is flat over \(S_f\). The
  closure-of-strata property (iii) reduces to the corresponding property in
  the \(n=0\) case for \(\pi_* \F(p)\) on \(S\) at any \(p \ge N\) separating all
  Hilbert polynomials.
\end{proof}

\subsection{Lean-side encodings of the sub-lemmas}
\label{sec:flatstrat_lean_helpers}

The three sub-lemmas above are realised in Lean by the following helper
declarations, which carry the substantive type signatures consumed by the
existence theorem. Each is an internal project helper with no external source
beyond Nitsure~\S4 already cited above.

\begin{lemma}

\leanok
[Flat-locus stratification, special case \(n = 0\)]
  \label{lem:flat_locus_stratification_lean}
  \lean{AlgebraicGeometry.flatLocusStratification}
  \uses{def:coherent_sheaf_flat}
  Let \(S\) be a noetherian scheme and \(\F\) a coherent \(\OO_S\)-module.
  Then there is a family of immersions \(\iota_e : S_e \to S\) indexed by
  \(e \in \mathbb{N}\) whose images are pairwise disjoint and set-theoretically
  cover \(|S|\), and such that for each \(e\) the pullback of \(\F\) along
  \(\iota_e\) is flat over \(S_e\) in the sense of
  \cref{def:coherent_sheaf_flat}.  This is the Lean encoding of the special
  case \(n = 0\) of the flattening stratification
  (\cref{lem:flat_locus_open}), the rank index \(e\) playing the role of
  \(\dim_{\kappa(s)} \F|_s\); the statement itself does not carry the rank
  labelling, so any \(\mathbb{N}\)-indexed disjoint flat family of immersions
  witnesses it.
\end{lemma}

\begin{proof}
  \leanok
  \uses{thm:flattening_stratification_exists}
  Apply the existence theorem \cref{thm:flattening_stratification_exists} to
  the identity morphism \(\pi = \mathrm{id}_S\) and \(\F\): it yields a
  finite set \(I\), immersions \(\iota_f : V_f \to S\) with pairwise
  disjoint images covering \(|S|\), and flatness of the pullback of \(\F\)
  to \(S \times_S V_f\) over \(V_f\).  The second projection
  \(p_f : S \times_S V_f \to V_f\) of the fibre product along the identity
  is an isomorphism, and the first projection factors as
  \(\mathrm{pr}_1 = p_f \circ \iota_f\); take \(S \times_S V_f\) itself
  as the stratum with \(\mathrm{pr}_1\) as its immersion, so disjointness
  and covering of images transfer along the surjection \(p_f\).  Flatness of
  the pullback over the stratum itself follows because the predicate of
  \cref{def:coherent_sheaf_flat} is insensitive to composing the base leg
  with an isomorphism: an affine open of the target corresponds under the
  isomorphism to an affine open of the source, the two section-ring actions
  on an unchanged section module differ by composition with a bijective ring
  map, and flatness passes along such a factorization since the target ring
  is a free module of rank one over the source.  Finally choose a bijection
  of \(I\) with an initial segment of \(\mathbb{N}\) and pad the family at
  the remaining indices by the empty scheme: its section rings are trivial,
  so every section module over them is flat, and the empty image preserves
  disjointness and covering.
\end{proof}

\subsection{Flatness transfer to strata}
\label{subsec:flat_transfer_strata}

The Noetherian-induction reduction is assembled from four transfer lemmas
(run 0010, T12 r6): flatness under pushout base change, the affine piece
analysis of a fibre product, affine-locality of flatness over the base, and
the flat stratum over one irreducible component.

\begin{lemma}[Flatness under pushout base change]

\leanok
  \label{lem:flat_pushout_base_change}
  \lean{Module.Flat.of_isPushout}
  % Module form of Stacks 00HI (3): flatness is stable under arbitrary base
  % change. Mathlib has the special case A = R, B = S
  % (\texttt{Module.Flat.baseChange}) and the ring form
  % (\texttt{RingHom.Flat.isStableUnderBaseChange}).
  Let \(R \to S\) and \(R \to A\) be commutative ring maps with pushout
  \(B = S \otimes_R A\), let \(M\) be an \(A\)-module which is flat over
  \(R\), and let \(N\) be a \(B\)-module which is a base change of \(M\)
  along \(A \to B\), i.e.\ \(N \cong B \otimes_A M\) as \(B\)-modules.  Then
  \(N\) is flat over \(S\).
\end{lemma}

\begin{proof}
  \leanok
  It suffices to prove that \(B \otimes_A M\) is flat over \(S\), and for
  this we show that the canonical \(R\)-linear map
  \(\eta : M \to B \otimes_A M\), \(m \mapsto 1 \otimes m\), exhibits
  \(B \otimes_A M\) as the base change \(S \otimes_R M\); flatness over
  \(S\) then follows from flatness of \(M\) over \(R\) since base change
  along \(R \to S\) preserves flatness.  We verify the universal property:
  let \(Q\) be an \(S\)-module and \(g : M \to Q\) an \(R\)-linear map.  The
  family \(\beta : A \to \operatorname{Hom}_R(M, Q)\),
  \(\beta(a)(m) = g(am)\), is \(R\)-linear into an \(S\)-module, so by the
  universal property of the ring base change \(B = S \otimes_R A\) it lifts
  to an \(S\)-linear map \(\Lambda : B \to \operatorname{Hom}_R(M, Q)\) with
  \(\Lambda(1 \otimes a) = \beta(a)\).  The pairing
  \((b, m) \mapsto \Lambda(b)(m)\) is additive, \(S\)-linear in \(b\), and
  \(A\)-balanced: on generators \(b = s (1 \otimes a')\) of \(B\) one has
  \(\Lambda((1 \otimes a) b)(m) = s\, g(a a' m) = \Lambda(b)(a m)\), and the
  general case follows since such elements generate \(B\) additively.  Hence
  the pairing descends to an \(S\)-linear map
  \(h : B \otimes_A M \to Q\) with \(h \circ \eta = g\).  Uniqueness: any
  \(S\)-linear \(h'\) with \(h' \circ \eta = g\) agrees with \(h\) on
  elements \((s(1 \otimes a)) \otimes m = s\,(1 \otimes am)\), hence on all
  of \(B \otimes_A M\).
\end{proof}

\begin{lemma}[Flat sections on an affine piece of a fibre product]

\leanok
  \label{lem:flat_section_pullback_piece}
  \lean{AlgebraicGeometry.flat_section_pullback_piece}
  \uses{def:coherent_sheaf_flat, lem:flat_pushout_base_change,
    def:pullback_app_isoTensor_sigma}
  Let
  \[
    \begin{array}{ccc}
      Y & \xrightarrow{\;g\;} & X \\
      \downarrow{\scriptstyle i_Y} & & \downarrow{\scriptstyle i_X} \\
      T & \xrightarrow{\;f\;} & S
    \end{array}
  \]
  be a cartesian square of schemes, and let \(U_S \subseteq S\),
  \(U_X \subseteq i_X^{-1}(U_S)\), \(U_T \subseteq f^{-1}(U_S)\) be affine
  opens.  Consider the \emph{piece}
  \(U_Y := g^{-1}(U_X) \cap i_Y^{-1}(U_T) \subseteq Y\).  Then \(U_Y\) is an
  affine open, the section-ring square
  \(\Gamma(S, U_S) \to \Gamma(X, U_X),\ \Gamma(S, U_S) \to \Gamma(T, U_T)\)
  with corner \(\Gamma(Y, U_Y)\) is a pushout of commutative rings, and for
  every quasi-coherent \(\OO_X\)-module \(\F\) such that
  \(\Gamma(\F, U_X)\) is flat over \(\Gamma(S, U_S)\), the sections
  \(\Gamma(g^* \F, U_Y)\) are flat over \(\Gamma(T, U_T)\).  If moreover
  \(\Gamma(\F, U_X)\) is a finite \(\Gamma(X, U_X)\)-module, then
  \(\Gamma(g^* \F, U_Y)\) is a finite \(\Gamma(Y, U_Y)\)-module.
\end{lemma}

\begin{proof}
  \leanok
  Restricting the cartesian square over the affine cospan
  \(U_X \to U_S \leftarrow U_T\) exhibits \(U_Y\) as the fibre product
  \(U_X \times_{U_S} U_T\) of affine schemes
  (Mathlib: \texttt{Scheme.Hom.isPullback\_resLE}), which is affine; applying
  global sections to this pullback of affines and reflecting along
  \(\Spec\) yields the pushout of section rings (Mathlib:
  \texttt{AlgebraicGeometry.isIso\_pushoutSection\_of\_isAffineOpen}).
  Since \(U_Y \subseteq g^{-1}(U_X)\) with \(U_X\), \(U_Y\) affine and
  \(\F\) quasi-coherent, the base-change section formula
  (\cref{def:pullback_app_isoTensor_sigma}, Stacks 01HQ/01I8) identifies
  \(\Gamma(g^* \F, U_Y) \cong
    \Gamma(Y, U_Y) \otimes_{\Gamma(X, U_X)} \Gamma(\F, U_X)\)
  as \(\Gamma(Y, U_Y)\)-modules, i.e.\ exhibits \(\Gamma(g^*\F, U_Y)\) as a
  base change of \(\Gamma(\F, U_X)\) along
  \(\Gamma(X, U_X) \to \Gamma(Y, U_Y)\).  Flatness over \(\Gamma(T, U_T)\)
  is now \cref{lem:flat_pushout_base_change} applied to the section-ring
  pushout square; finiteness is preserved because a base change of a finite
  module is finite.
\end{proof}

\begin{lemma}[Flatness over the base is affine-local]

\leanok
  \label{lem:flat_affine_local}
  \lean{AlgebraicGeometry.flat_section_of_affine_cover}
  \uses{def:coherent_sheaf_flat, lem:appLE_base_res_apply}
  Let \(q : Y \to T\) be a morphism of schemes and \(\mathcal G\) a quasi-coherent
  \(\OO_Y\)-module.  Suppose given affine opens
  \(U_j \subseteq T\) and \(W_j \subseteq q^{-1}(U_j)\) (\(j \in J\)) such
  that the \(W_j\) cover \(Y\) and each \(\Gamma(\mathcal G, W_j)\) is flat over
  \(\Gamma(T, U_j)\).  Then for \emph{every} pair of affine opens
  \(U \subseteq T\), \(W \subseteq q^{-1}(U)\), the sections
  \(\Gamma(\mathcal G, W)\) are flat over \(\Gamma(T, U)\).
\end{lemma}

\begin{proof}
  \leanok
  This is the two-layer basic-open reduction of
  \cref{lem:flat_section_pair}, with the freeness input replaced by chart
  flatness.  Fix an affine pair \((U, W)\).  For each \(x \in W\) choose a
  chart \(W_j \ni x\); since \(q(x) \in U \cap U_j\) and both are affine,
  there is a basic open \(D(a) = D(a')\) of \(T\) common to \(U\) and
  \(U_j\) containing \(q(x)\), and then a basic open
  \(D(b) = D(c) \ni x\) common to \(W\) and \(W_j\) with
  \(D(b) \subseteq q^{-1}(D(a))\).  The elements \(b\) generate the unit
  ideal of \(\Gamma(Y, W)\), and each restriction
  \(\Gamma(\mathcal G, W) \to \Gamma(\mathcal G, D(b))\) is a localization at the powers of
  \(b\) (\cref{lem:qcoh_section_localization_basicOpen}), so flatness of
  \(\Gamma(\mathcal G, W)\) over \(\Gamma(T, U)\) may be checked on the pieces
  \(\Gamma(\mathcal G, D(b))\).  On a piece: flatness of \(\Gamma(\mathcal G, W_j)\) over
  \(\Gamma(T, U_j)\) passes to the fibre-side localization
  \(\Gamma(\mathcal G, D(c))\) (mixed-base stability of flatness under localization
  of the module at a submonoid of the fibre algebra), and then the base ring
  is exchanged twice along
  \(\Gamma(T, U_j) \to \Gamma(T, D(a')) = \Gamma(T, D(a))\) and
  \(\Gamma(T, U) \to \Gamma(T, D(a))\) (for a module over a localized ring,
  flatness over the ring and over its localization agree; the compatibility
  of the induced module structures is \cref{lem:appLE_base_res_apply}).
\end{proof}

\begin{lemma}[Reduced subscheme of an irreducible closed subset is integral]

\leanok
  \label{lem:vanishing_ideal_subscheme_integral}
  \lean{AlgebraicGeometry.isIntegral_vanishingIdeal_subscheme}
  Let \(S\) be a scheme and \(Z \subseteq S\) a closed subset.  The
  subscheme cut out by the vanishing ideal sheaf of \(Z\) (the reduced
  induced structure) is reduced; if \(Z\) is irreducible (in particular
  non-empty), it is integral.
\end{lemma}

\begin{proof}
  \leanok
  The vanishing ideal sheaf is radical: it equals the vanishing ideal of
  its own support (Galois connection between supports and vanishing
  ideals), which is the radical.  The subscheme is covered by the affine
  charts \(\Spec(\Gamma(S, U)/I(U))\) for \(U\) ranging over affine opens
  of \(S\), and each \(I(U)\) is a radical ideal, so each chart is the
  spectrum of a reduced ring; hence the subscheme is reduced.  Its
  underlying space is the subspace \(Z\), which is irreducible by
  hypothesis, and a reduced scheme with irreducible non-empty underlying
  space is integral.
\end{proof}

\begin{lemma}[Flat stratum over an irreducible component]

\leanok
  \label{lem:flat_stratum_component}
  \lean{AlgebraicGeometry.flat_stratum_of_irreducible}
  \uses{def:coherent_sheaf_flat, thm:generic_flatness,
    lem:vanishing_ideal_subscheme_integral, lem:flat_section_pullback_piece,
    lem:flat_affine_local}
  Let \(S\) be a noetherian scheme, \(\pi : X \to S\) proper, \(\F\) a
  coherent \(\OO_X\)-module, and \(Z_c \subseteq S\) an irreducible closed
  subset with reduced subscheme \(T \to S\).  Then there is a non-empty open
  \(\Omega \subseteq T\) such that for \emph{every} open
  \(\mathit{St} \subseteq \Omega\), regarded as a locally closed subscheme
  \(\mathit{St} \to S\), the pullback of \(\F\) to
  \(X \times_S \mathit{St}\) is flat over \(\mathit{St}\) in the sense of
  \cref{def:coherent_sheaf_flat}.
\end{lemma}

\begin{proof}
  \leanok
  \(T\) is integral (\cref{lem:vanishing_ideal_subscheme_integral}) and
  noetherian (a closed subscheme of a noetherian scheme), and the base
  change \(p : X \times_S T \to T\) is proper, hence quasi-compact and
  locally of finite type.  The pullback of \(\F\) to \(X \times_S T\) is
  quasi-coherent, and it admits affine charts with finite section modules:
  around any point, choose affine opens \(U_S \subseteq S\),
  \(U_T \subseteq T\) above it, and an affine chart \(W_X \subseteq X\)
  with finite sections (Stacks 01PC); the corresponding piece is an affine
  chart of \(X \times_S T\) whose sections are the base change of
  \(\Gamma(\F, W_X)\), hence finite (\cref{lem:flat_section_pullback_piece}),
  and this finiteness localizes to a basic open inside any given
  neighbourhood.  Generic flatness (\cref{thm:generic_flatness}, in its
  quasi-coherent form with this chart supply) yields a non-empty open
  \(\Omega \subseteq T\) with flat sections on all affine pairs below
  \(\Omega\).

  Now let \(\mathit{St} \subseteq \Omega\) be open.  The stratum family
  \(X \times_S \mathit{St} \to \mathit{St}\) maps to the component family by
  the comparison morphism
  \(\kappa : X \times_S \mathit{St} \to X \times_S T\), and the square
  formed by \(\kappa\) and the two projections over
  \(\mathit{St} \hookrightarrow T\) is cartesian (cancellation of pasted
  pullback squares).  Cover \(X \times_S \mathit{St}\) by pieces
  \(g^{-1}(W_X) \cap q^{-1}(U')\) with \(U' \subseteq \mathit{St}\) affine
  lying over an affine \(U_1 \subseteq \Omega\); on each such piece,
  applying \cref{lem:flat_section_pullback_piece} to the cartesian
  \(\kappa\)-square with flatness input from the generic-flatness output on
  the corresponding piece of \(X \times_S T\) (an affine open below
  \(\Omega\)) yields flat sections over \(U'\).  By affine-locality
  (\cref{lem:flat_affine_local}) all affine pairs of the stratum family
  have flat sections; finally the pulled-back sheaf along \(\kappa\)
  composed with the projection is identified with the pullback of \(\F\)
  along the projection of the stratum family (functoriality of pullback,
  \(g \circ \kappa = g'\)), so the flatness predicate transports along this
  isomorphism of module sheaves.
\end{proof}

\begin{lemma}

\leanok
[Noetherian-induction reduction to the flat case]
  \label{lem:flat_locus_reduction_lean}
  \lean{AlgebraicGeometry.flatLocusReduction}
  \uses{def:coherent_sheaf_flat, thm:generic_flatness}
  Let \(S\) be a noetherian scheme, \(\pi : X \to S\) a proper morphism,
  and \(\F\) a coherent \(\OO_X\)-module. Then there exist finitely many
  immersions \(\iota_i : V_i \to S\) with pairwise disjoint images
  set-theoretically covering \(|S|\), such that for each \(i\) the pullback of
  \(\F\) to \(X \times_S V_i\) is flat over \(V_i\) in the sense of
  \cref{def:coherent_sheaf_flat}.  This is the Lean encoding of the
  Noetherian-induction reduction (\cref{lem:nonflat_locus_proper}), obtained by
  peeling off non-empty flat open patches via generic flatness
  (\cref{thm:generic_flatness}).

  Noetherianity of \(S\) (rather than mere local noetherianity) is necessary
  for the \emph{finiteness} of the family: on
  \(S = \coprod_{n \ge 1} \mathbb{A}^n\), equip the \(n\)-th component with
  the coherent sheaf \(\bigoplus_{k \le n} (i_{V_k})_* \OO_{V_k}\) for a
  strictly nested flag \(V_1 \supset \cdots \supset V_n\) of integral closed
  subvarieties.  A locally-closed stratum on which this sheaf is flat (hence
  locally free) meets the flag levels in relatively clopen subsets, so a
  stratum through the generic point of \(V_k\) cannot also contain the
  generic point of \(V_{k'}\) for \(k' \neq k\): each neighbourhood of the
  latter meets the dense lower level.  The \(n\)-th component therefore
  requires at least \(n + 1\) strata, and no finite family covers all
  components.
\end{lemma}

\begin{proof}
  \leanok
  \uses{lem:nonflat_locus_proper, thm:generic_flatness,
    lem:flat_stratum_component, lem:vanishing_ideal_subscheme_integral}
  By well-founded induction on the closed subsets of \(S\) (the noetherian
  hypothesis makes strict inclusion of closed subsets well-founded), we prove:
  for every closed \(Z \subseteq S\) there is a finite family of immersions
  with pairwise disjoint images whose union is exactly \(Z\) and with flat
  pullback of \(\F\) on each member.  For \(Z = \emptyset\) take the empty
  family.

  For \(Z \neq \emptyset\), let \(Z_1, \dots, Z_r \subseteq S\) be the images
  of the (finitely many, by noetherianity) irreducible components of the
  subspace \(Z\); each \(Z_c\) is an irreducible closed subset of \(S\)
  contained in \(Z\), and together they cover \(Z\).
  \Cref{lem:flat_stratum_component} applied to the reduced subscheme
  \(T_c \to S\) of \(Z_c\) yields a non-empty open \(\Omega_c \subseteq T_c\)
  all of whose open subsets are flat strata.  Let
  \(E_c \subseteq S\) be the open complement of the union of the other
  components, and take as stratum the open
  \(\mathit{St}_c := \Omega_c \cap (T_c \to S)^{-1}(E_c) \subseteq T_c\),
  a locally closed subscheme of \(S\) via
  \(\mathit{St}_c \hookrightarrow T_c \to S\), with flat pullback by the
  choice of \(\Omega_c\).

  The images of the \(\mathit{St}_c\) are pairwise disjoint: the image of
  \(\mathit{St}_c\) is contained in \(Z_c \setminus \bigcup_{c' \neq c} Z_{c'}\).
  Their union \(V\) is relatively open in \(Z\): realizing the open image of
  each \(\mathit{St}_c\) inside \(Z_c\) by an open \(W_c \subseteq S\), one has
  \(V = Z \cap \bigcup_c (W_c \cap E_c)\), because a point of
  \(Z \cap W_c \cap E_c\) avoids all components other than \(Z_c\), hence lies
  in \(Z_c\), hence in the image of \(\mathit{St}_c\).  Moreover \(V\) is
  non-empty: for any component \(Z_{c}\), irredundancy of the component
  decomposition shows \(Z_c \not\subseteq \bigcup_{c' \neq c} Z_{c'}\), so the
  open \((T_c \to S)^{-1}(E_c)\) is non-empty, and two non-empty opens of the
  irreducible space \(T_c\) intersect, so \(\mathit{St}_c \neq \emptyset\).

  Therefore \(Z' := Z \setminus V\) is a closed subset strictly smaller than
  \(Z\); the induction hypothesis provides a finite disjoint flat family with
  union \(Z'\), and adjoining the strata \(\mathit{St}_c\) gives the required
  family for \(Z\) (disjointness across the two groups holds because the
  recursive family lies in \(Z'\), which is disjoint from \(V\)).  Applying
  the statement to \(Z = S\) proves the lemma.
\end{proof}

\begin{lemma}

\leanok
[Assembly of the Hilbert-polynomial-indexed strata]
  \label{lem:flat_locus_assembly_lean}
  \lean{AlgebraicGeometry.flatLocusAssembly}
  \uses{def:coherent_sheaf_flat}
  Let \(S\) be a noetherian scheme, \(\pi : X \to S\) a proper morphism,
  and \(\F\) a coherent \(\OO_X\)-module. Then there exist a finite index set
  \(I\), immersions \(\iota_f : S_f \to S\) (\(f \in I\)) with pairwise disjoint
  images set-theoretically covering \(|S|\), and an injective label map
  \(P : I \to (\mathbb{N} \to \Z)\) (so \(f = g \iff P_f = P_g\)), such that for
  each \(f\) the pullback of \(\F\) to \(X \times_S S_f\) is flat over \(S_f\) in
  the sense of \cref{def:coherent_sheaf_flat}.  In the intended refinement
  (\cref{lem:noetherian_induction_strata}) the label \(P_f\) is the Hilbert
  polynomial of the fibres over \(S_f\); the present statement records only
  the injective labelling of the strata.
\end{lemma}

\begin{proof}
  \leanok
  \uses{lem:flat_locus_reduction_lean}
  Since the statement requires of \(P\) only injectivity, it follows from the
  reduction \cref{lem:flat_locus_reduction_lean}: take its stratification and
  label the finitely many strata by any injection
  \(I \hookrightarrow \mathbb{N} \to \Z\) (for instance constant functions
  with pairwise distinct values).  The genuinely Hilbert-polynomial-indexed
  refinement is the content of \cref{lem:noetherian_induction_strata}.
\end{proof}

\section{The entry ideal of a matrix presentation}
\label{sec:entry_ideal}

The rank strata of \cref{lem:flat_locus_open} carry a canonical --- in
general non-reduced --- scheme structure, cut out locally by the ideal of
entries of the relation matrix of a presentation.  This section develops the
commutative-algebra substrate: presentations of a module by \(e\)
generators, the entry ideal, its base-change behaviour, Nitsure's local
universal property, independence of the chosen presentation, and the
Nakayama prolongation producing \(e\)-generator presentations near a point
of fiber rank \(e\).  Throughout, \(R\) is a commutative ring and \(M\) an
\(R\)-module.

\begin{definition}[Matrix presentation]
  \label{def:matrix_presentation}
  \lean{Module.MatrixPresentation}
  \leanok
  \source{nitsure-hilbert-quot}
  A \emph{matrix presentation} of \(M\) by \(e\) generators and \(m\)
  relations consists of a matrix
  \(\psi \in \operatorname{Mat}_{e \times m}(R)\) and a surjection
  \(\pi : R^{\oplus e} \to M\) making the sequence
  \[
    R^{\oplus m} \xrightarrow{\ \psi\ } R^{\oplus e}
      \xrightarrow{\ \pi\ } M \to 0
  \]
  exact.
\end{definition}

\begin{lemma}[Finite presentation]
  \label{lem:matrix_presentation_fp}
  \lean{Module.MatrixPresentation.finitePresentation}
  \leanok
  \uses{def:matrix_presentation}
  A module admitting a matrix presentation is finitely presented.
\end{lemma}

\begin{proof}
  \leanok
  \(\pi\) is a surjection from a finite free module whose kernel, the image
  of \(\psi\), is generated by the \(m\) columns of \(\psi\), hence finitely
  generated.
\end{proof}

\begin{definition}[Entry ideal]
  \label{def:entry_ideal}
  \lean{Module.MatrixPresentation.entryIdeal}
  \leanok
  \uses{def:matrix_presentation}
  \source{nitsure-hilbert-quot}
  The \emph{entry ideal} \(I(\psi) \subseteq R\) of a matrix presentation is
  the ideal generated by the \(e \cdot m\) entries \(\psi_{ij}\) of the
  relation matrix.  It is the ideal of \(1 \times 1\) minors of \(\psi\);
  for an \(e\)-generator presentation this is the Fitting ideal
  \(\operatorname{Fitt}_{e-1}(M)\).
\end{definition}

\begin{lemma}[Vanishing criterion]
  \label{lem:entry_ideal_vanish}
  \lean{Module.MatrixPresentation.relMatrix_map_eq_zero_iff}
  \leanok
  \uses{def:entry_ideal}
  For an \(R\)-algebra \(A\), the entrywise image matrix
  \(\psi \otimes A \in \operatorname{Mat}_{e \times m}(A)\) vanishes if and
  only if \(I(\psi) \subseteq \ker(R \to A)\).
\end{lemma}

\begin{proof}
  \leanok
  Both sides assert that every entry of \(\psi\) maps to \(0\) in \(A\).
\end{proof}

\begin{definition}[Base change of a presentation]
  \label{def:matrix_presentation_base_change}
  \lean{Module.MatrixPresentation.baseChange}
  \leanok
  \uses{def:matrix_presentation}
  \source{nitsure-hilbert-quot}
  Let \(A\) be an \(R\)-algebra.  The matrix \(\psi \otimes A\) obtained by
  applying \(R \to A\) entrywise, together with the tensored projection
  \(A^{\oplus e} \cong A \otimes_R R^{\oplus e}
    \xrightarrow{1 \otimes \pi} A \otimes_R M\),
  is a matrix presentation of \(A \otimes_R M\) by \(e\) generators and
  \(m\) relations: the tensored sequence
  \(A^{\oplus m} \to A^{\oplus e} \to A \otimes_R M \to 0\) remains exact by
  right-exactness of the tensor product.
\end{definition}

\begin{lemma}[Entry ideal under base change]
  \label{lem:entry_ideal_base_change}
  \lean{Module.MatrixPresentation.entryIdeal_baseChange}
  \leanok
  \uses{def:entry_ideal, def:matrix_presentation_base_change}
  The entry ideal of the base-changed presentation is the image ideal:
  \(I(\psi \otimes A) = I(\psi) \cdot A\).
\end{lemma}

\begin{proof}
  \leanok
  The generators of \(I(\psi \otimes A)\) are exactly the images of the
  generators of \(I(\psi)\).
\end{proof}

\begin{lemma}[Vanishing relations give an isomorphism]
  \label{lem:presentation_iso_of_vanish}
  \lean{Module.MatrixPresentation.projEquiv}
  \leanok
  \uses{def:matrix_presentation}
  If \(\psi = 0\), then \(\pi : R^{\oplus e} \to M\) is an isomorphism.
\end{lemma}

\begin{proof}
  \leanok
  \(\ker \pi = \operatorname{im} \psi = 0\) by exactness, and \(\pi\) is
  surjective.
\end{proof}

\begin{lemma}[Vanishing relations give a free module]
  \label{lem:presentation_free_of_vanish}
  \lean{Module.MatrixPresentation.free_of_relMatrix_eq_zero}
  \leanok
  \uses{lem:presentation_iso_of_vanish}
  If \(\psi = 0\), then \(M\) is free.
\end{lemma}

\begin{proof}
  \leanok
  Transport freeness along the isomorphism of
  \cref{lem:presentation_iso_of_vanish}.
\end{proof}

\begin{lemma}[Vanishing relations give constant rank]
  \label{lem:presentation_rank_of_vanish}
  \lean{Module.MatrixPresentation.rankAtStalk_eq_of_relMatrix_eq_zero}
  \leanok
  \uses{lem:presentation_iso_of_vanish}
  If \(\psi = 0\), then the localization \(M_{\mathfrak p}\) is free of rank
  \(e\) over \(R_{\mathfrak p}\) for every prime
  \(\mathfrak p \subseteq R\).
\end{lemma}

\begin{proof}
  \leanok
  By \cref{lem:presentation_iso_of_vanish}, \(M \cong R^{\oplus e}\); a ring
  with a prime ideal is nontrivial, so \(R^{\oplus e}\) has constant rank
  \(e\) at every prime.
\end{proof}

\begin{theorem}[Flat everywhere-rank-\(e\) modules have vanishing relation
  matrix]
  \label{thm:entry_ideal_flat_vanish}
  \lean{Module.MatrixPresentation.relMatrix_eq_zero_of_flat}
  \leanok
  \uses{def:matrix_presentation}
  \source{nitsure-hilbert-quot}
  Let \(M\) carry a matrix presentation \((\psi, \pi)\) by \(e\) generators.
  If \(M\) is flat and \(M_{\mathfrak p}\) has rank \(e\) over
  \(R_{\mathfrak p}\) for every prime \(\mathfrak p\), then \(\psi = 0\).
  (This is the direction of Nitsure's local universal property left implicit
  in the source: a surjection \(R^{\oplus e} \twoheadrightarrow M\) onto a
  locally free module of everywhere-rank \(e\) has zero kernel.)
\end{theorem}

\begin{proof}
  \leanok
  \uses{lem:matrix_presentation_fp}
  By \cref{lem:matrix_presentation_fp}, \(M\) is finitely presented, hence
  --- being flat --- projective.  The surjection \(\pi\) therefore splits:
  choose \(s : M \to R^{\oplus e}\) with \(\pi \circ s = \mathrm{id}\).
  The splitting yields an isomorphism
  \(R^{\oplus e} \cong M \times \ker \pi\), \(x \mapsto
  (\pi x,\, x - s(\pi x))\), so \(K := \ker \pi\) is a direct summand of a
  finite free module, hence itself finite projective (in particular flat).
  Localizing at any prime \(\mathfrak p\) and counting ranks,
  \(e = \operatorname{rk}_{\mathfrak p} R^{\oplus e}
     = \operatorname{rk}_{\mathfrak p} M + \operatorname{rk}_{\mathfrak p} K
     = e + \operatorname{rk}_{\mathfrak p} K\),
  so \(K\) has rank \(0\) at every prime; a finite flat module of
  everywhere-rank \(0\) is zero.  Hence
  \(\operatorname{im} \psi = \ker \pi = 0\) by exactness, and every entry
  \(\psi_{ij} = (\psi(\delta_j))_i\) vanishes.
\end{proof}

\begin{lemma}[Comparison of entry ideals]
  \label{lem:entry_ideal_le}
  \lean{Module.MatrixPresentation.entryIdeal_le_entryIdeal}
  \leanok
  \uses{def:entry_ideal}
  If \((\psi, \pi)\) and \((\psi', \pi')\) are two matrix presentations of
  \(M\) by \(e\) generators (with possibly different numbers of relations),
  then \(I(\psi') \subseteq I(\psi)\).
\end{lemma}

\begin{proof}
  \leanok
  \uses{def:matrix_presentation_base_change, lem:entry_ideal_vanish,
    lem:presentation_free_of_vanish, lem:presentation_rank_of_vanish,
    thm:entry_ideal_flat_vanish}
  Base change both presentations to \(A := R / I(\psi)\).  The entries of
  \(\psi \otimes A\) vanish by construction
  (\cref{lem:entry_ideal_vanish}), so by
  \cref{lem:presentation_free_of_vanish,lem:presentation_rank_of_vanish}
  the module \(A \otimes_R M\) is free (hence flat) of rank \(e\) at every
  prime of \(A\).  Applying \cref{thm:entry_ideal_flat_vanish} to the
  base-changed presentation \((\psi' \otimes A, \pi' \otimes A)\) gives
  \(\psi' \otimes A = 0\), i.e.\ every entry of \(\psi'\) lies in
  \(\ker(R \to R/I(\psi)) = I(\psi)\).
\end{proof}

\begin{theorem}[Presentation-independence of the entry ideal]
  \label{thm:entry_ideal_presentation_independent}
  \lean{Module.MatrixPresentation.entryIdeal_eq}
  \leanok
  \uses{def:entry_ideal}
  \source{nitsure-hilbert-quot}
  Any two matrix presentations of \(M\) by \(e\) generators have the same
  entry ideal.  (This is the presentation-independence of the Fitting ideal
  \(\operatorname{Fitt}_{e-1}\), and is what lets the local entry-ideal
  subschemes of \cref{lem:flat_locus_open} glue.)
\end{theorem}

\begin{proof}
  \leanok
  \uses{lem:entry_ideal_le}
  Apply \cref{lem:entry_ideal_le} in both directions.
\end{proof}

\begin{definition}[Fiber rank]
  \label{def:fiber_rank}
  \lean{Ideal.fiberRank}
  \leanok
  \source{nitsure-hilbert-quot}
  For a prime \(\mathfrak p \subseteq R\) with residue field
  \(\kappa(\mathfrak p)\), the \emph{fiber rank} of \(M\) at
  \(\mathfrak p\) is
  \(\dim_{\kappa(\mathfrak p)} \bigl(\kappa(\mathfrak p) \otimes_R M\bigr)\).
\end{definition}

\begin{theorem}[Point-set of the entry-ideal locus]
  \label{thm:entry_ideal_prime_iff}
  \lean{Module.MatrixPresentation.entryIdeal_le_prime_iff}
  \leanok
  \uses{def:entry_ideal, def:fiber_rank}
  \source{nitsure-hilbert-quot}
  Let \((\psi, \pi)\) be a matrix presentation of \(M\) by \(e\) generators
  and \(\mathfrak p \subseteq R\) a prime.  Then
  \(I(\psi) \subseteq \mathfrak p\) if and only if the fiber rank of \(M\)
  at \(\mathfrak p\) equals \(e\).
\end{theorem}

\begin{proof}
  \leanok
  \uses{def:matrix_presentation_base_change, lem:entry_ideal_vanish,
    lem:presentation_iso_of_vanish, thm:entry_ideal_flat_vanish}
  Base change the presentation to the residue field
  \(\kappa(\mathfrak p)\), whose structure map has kernel exactly
  \(\mathfrak p\).  If \(I(\psi) \subseteq \mathfrak p\), the fiber
  presentation has vanishing relation matrix
  (\cref{lem:entry_ideal_vanish}), so the fiber is isomorphic to
  \(\kappa(\mathfrak p)^{\oplus e}\)
  (\cref{lem:presentation_iso_of_vanish}) and has dimension \(e\).
  Conversely, if the fiber has dimension \(e\), then --- every module over
  a field being free, and free modules over a field having rank equal to
  their dimension at the unique relevant localization --- the fiber
  presentation satisfies the hypotheses of
  \cref{thm:entry_ideal_flat_vanish}, so its relation matrix vanishes and
  every entry of \(\psi\) lies in
  \(\ker(R \to \kappa(\mathfrak p)) = \mathfrak p\).
\end{proof}

\begin{theorem}[Nakayama prolongation of a fiber basis]
  \label{thm:nakayama_prolongation}
  \lean{Module.FinitePresentation.exists_matrixPresentation_of_isLocalizedModule}
  \leanok
  \uses{def:matrix_presentation, def:fiber_rank}
  \source{nitsure-hilbert-quot}
  Let \(M\) be finitely presented over \(R\), let \(\mathfrak p \subseteq R\)
  be a prime, and let \(e\) be the fiber rank of \(M\) at \(\mathfrak p\).
  Then there exists \(g \notin \mathfrak p\) such that every localization of
  \(M\) away from \(g\) --- that is, every module \(N\) over an \(R\)-algebra
  \(R'\) realizing the localizations of \(R\) and \(M\) at the powers of
  \(g\) --- admits a matrix presentation by exactly \(e\) generators.
\end{theorem}

\begin{proof}
  \leanok
  \uses{lem:matrix_presentation_fp}
  Every element of the fiber \(\kappa(\mathfrak p) \otimes_R M\) becomes,
  after scaling by some \(r \notin \mathfrak p\) (a unit of
  \(\kappa(\mathfrak p)\)), a pure tensor \(1 \otimes s\) with \(s \in M\):
  this holds for pure tensors \(c \otimes s\) because
  \(\kappa(\mathfrak p)\) is a localization of \(R/\mathfrak p\)-fractions
  (clear the denominator of \(c\)), and sums are handled by scaling both
  summands by the product of their two scalars.  Scaling each vector of a
  basis of the fiber this way produces \(m_1, \dots, m_e \in M\) whose
  images \(1 \otimes m_i\) again form a basis of the fiber (they differ from
  the original basis by unit scalars).

  Let \(\alpha : R^{\oplus e} \to M\) send the standard basis to the
  \(m_i\), and let \(Q := M / \operatorname{im} \alpha\); it is finitely
  generated.  Since the \(1 \otimes m_i\) span the fiber and the tensor
  product is right exact, \(\kappa(\mathfrak p) \otimes_R Q = 0\); as
  \(\kappa(\mathfrak p) \otimes_R Q\) is the fiber of the finite
  \(R_{\mathfrak p}\)-module \(Q_{\mathfrak p}\) at the maximal ideal,
  Nakayama gives \(Q_{\mathfrak p} = 0\).  Hence \(\mathfrak p\) lies
  outside the support of the finite module \(Q\), which is the zero locus
  of its annihilator, so there is \(g \in \operatorname{Ann}(Q)\) with
  \(g \notin \mathfrak p\); thus \(g \cdot M \subseteq
  \operatorname{im} \alpha\).

  Now let \(R'\) and \(N\) realize the localizations of \(R\) and \(M\)
  away from \(g\), with structure map \(l : M \to N\).  The images
  \(l(m_i)\) generate \(N\) over \(R'\): any \(n \in N\) satisfies
  \(g^k n = l(x)\) for some \(x \in M\) and some \(k\), the element
  \(g x\) lies in \(\operatorname{im} \alpha\), and \(g\) and \(g^k\) are
  units in \(R'\).  This gives a surjection
  \(\alpha' : R'^{\oplus e} \to N\).  Localization is a base change, so
  \(N\) is finitely presented over \(R'\)
  (\cref{lem:matrix_presentation_fp} for \(M\), transported along the base
  change); hence \(\ker \alpha'\) is finitely generated, and any finite
  generating family of \(\ker \alpha'\), arranged as the columns of a
  matrix, completes \(\alpha'\) to a matrix presentation of \(N\) by \(e\)
  generators.
\end{proof}

\section{The rank-\(e\) stratum as a closed subscheme}
\label{sec:flatstrat_rank_stratum}

The entry-ideal commutative algebra of \cref{sec:entry_ideal} is local to an
affine chart.  This section globalises it: for a quasi-coherent module
\(\mathcal G\) on a scheme \(X\) it assembles the local entry ideals into a
quasi-coherent ideal sheaf whose closed subscheme --- the \emph{rank-\(e\)
stratum} --- has support exactly the locus where the fibre of \(\mathcal G\)
has dimension \(e\).  This is the geometric substrate of Nitsure's
\(n = 0\) construction (\cref{lem:flat_locus_open}).  Throughout \(X\) is a
scheme and \(\mathcal G\) a quasi-coherent \(\OO_X\)-module; \(\Gamma(\mathcal G, U)\)
denotes its sections over an open \(U\), a module over \(\Gamma(X, U) = \OO_X(U)\).

\subsection{Base change of the fibre rank}

\begin{theorem}[Base-change invariance of the fibre rank]
  \label{thm:fiber_rank_base_change}
  \lean{Ideal.fiberRank_baseChange}
  \leanok
  \uses{def:fiber_rank}
  Let \(A\) be an \(R\)-algebra, \(M\) an \(R\)-module, and \(\mathfrak q\) a
  prime of \(A\) with contraction \(\mathfrak p = \mathfrak q \cap R\).  Then
  the fibre rank of \(A \otimes_R M\) at \(\mathfrak q\) equals the fibre rank
  of \(M\) at \(\mathfrak p\).  No flatness or finiteness hypotheses are
  needed.
\end{theorem}

\begin{proof}
  \leanok
  The residue-field map \(R \to \kappa(\mathfrak p) \to \kappa(\mathfrak q)\)
  makes \(\kappa(\mathfrak q)\) a \(\kappa(\mathfrak p)\)-algebra.  Cancelling
  base changes gives \(\kappa(\mathfrak p)\)-linear, then
  \(\kappa(\mathfrak q)\)-linear, isomorphisms
  \[
    \kappa(\mathfrak q) \otimes_A (A \otimes_R M)
      \;\cong\; \kappa(\mathfrak q) \otimes_R M
      \;\cong\; \kappa(\mathfrak q) \otimes_{\kappa(\mathfrak p)}
        \bigl(\kappa(\mathfrak p) \otimes_R M\bigr).
  \]
  Extension of scalars along a field extension preserves dimension, so
  \(\dim_{\kappa(\mathfrak q)}
    \kappa(\mathfrak q) \otimes_{\kappa(\mathfrak p)}
    (\kappa(\mathfrak p) \otimes_R M)
    = \dim_{\kappa(\mathfrak p)}(\kappa(\mathfrak p) \otimes_R M)\),
  which is the claim.
\end{proof}

\begin{lemma}[Fibre rank is a linear invariant]
  \label{lem:fiber_rank_congr}
  \lean{Ideal.fiberRank_congr}
  \leanok
  \uses{def:fiber_rank}
  If \(\sigma : M \cong N\) is an isomorphism of \(R\)-modules, then \(M\) and
  \(N\) have the same fibre rank at every prime \(\mathfrak p\).
\end{lemma}

\begin{proof}
  \leanok
  Base change \(\sigma\) along \(R \to \kappa(\mathfrak p)\) to obtain a
  \(\kappa(\mathfrak p)\)-linear isomorphism
  \(\kappa(\mathfrak p) \otimes_R M \cong \kappa(\mathfrak p) \otimes_R N\);
  isomorphic vector spaces have equal dimension.
\end{proof}

\begin{corollary}[Fibre rank of a localized module]
  \label{cor:fiber_rank_localized}
  \lean{Ideal.fiberRank_of_isLocalizedModule}
  \leanok
  \uses{thm:fiber_rank_base_change, lem:fiber_rank_congr}
  Let \(R' = S^{-1}R\) be a localization, \(N = S^{-1}M\) the corresponding
  localized module, and \(\mathfrak q\) a prime of \(R'\).  Then the fibre
  rank of \(N\) at \(\mathfrak q\) equals the fibre rank of \(M\) at the
  contraction \(\mathfrak q \cap R\).
\end{corollary}

\begin{proof}
  \leanok
  Localization is a base change: the canonical map \(M \to N\) induces an
  \(R'\)-linear isomorphism \(R' \otimes_R M \cong N\).  By
  \cref{lem:fiber_rank_congr} the fibre rank of \(N\) at \(\mathfrak q\)
  equals that of \(R' \otimes_R M\), which by \cref{thm:fiber_rank_base_change}
  equals the fibre rank of \(M\) at \(\mathfrak q \cap R\).
\end{proof}

\subsection{Presentation charts and the point rank}

\begin{definition}[Presentation chart]
  \label{def:presentation_chart}
  \lean{AlgebraicGeometry.Scheme.Modules.IsPresentationChart}
  \leanok
  \uses{def:matrix_presentation}
  \source{nitsure-hilbert-quot}
  An affine open \(V \subseteq X\) is an \emph{\(e\)-presentation chart} for
  \(\mathcal G\) if the section module \(\Gamma(\mathcal G, V)\) admits a
  matrix presentation by \(e\) generators over \(\Gamma(X, V)\), that is, a
  right-exact sequence
  \(\OO_V^{\oplus m} \xrightarrow{\psi} \OO_V^{\oplus e}
    \to \mathcal G|_V \to 0\).
\end{definition}

\begin{definition}[Restriction of a presentation to a basic open]
  \label{def:presentation_basic_open}
  \lean{AlgebraicGeometry.Scheme.Modules.MatrixPresentationBasicOpen}
  \leanok
  \uses{def:matrix_presentation_base_change, def:presentation_chart}
  Let \(P = (\psi, \pi)\) be a matrix presentation of
  \(\Gamma(\mathcal G, V)\) by \(e\) generators over an affine open \(V\), and
  \(f \in \Gamma(X, V)\).  Since \(\mathcal G\) is quasi-coherent, restriction
  exhibits \(\Gamma(\mathcal G, D(f))\) as the localization
  \(\Gamma(\mathcal G, V) \otimes_{\Gamma(X, V)} \Gamma(X, D(f))\).  Base
  change of \(P\) along \(\Gamma(X, V) \to \Gamma(X, D(f))\)
  (\cref{def:matrix_presentation_base_change}), transported along this
  localization isomorphism, is a matrix presentation of
  \(\Gamma(\mathcal G, D(f))\) by \(e\) generators, the \emph{restriction of
  \(P\) to \(D(f)\)}; its relation matrix is the entrywise restriction
  \(\psi|_{D(f)}\).
\end{definition}

\begin{lemma}[Entry ideal of a restricted presentation]
  \label{lem:entry_ideal_basic_open}
  \lean{AlgebraicGeometry.Scheme.Modules.entryIdeal_matrixPresentationBasicOpen}
  \leanok
  \uses{def:presentation_basic_open, lem:entry_ideal_base_change}
  With the notation of \cref{def:presentation_basic_open}, the entry ideal of
  the restricted presentation is the image of \(I(\psi)\) under restriction:
  \(I(\psi|_{D(f)}) = I(\psi) \cdot \Gamma(X, D(f))\).
\end{lemma}

\begin{proof}
  \leanok
  The restricted presentation is a base change of \(P\), so its entry ideal is
  the image ideal by \cref{lem:entry_ideal_base_change}.
\end{proof}

\begin{lemma}[Presentation charts pass to basic opens]
  \label{lem:presentation_chart_basic_open}
  \lean{AlgebraicGeometry.Scheme.Modules.IsPresentationChart.basicOpen}
  \leanok
  \uses{def:presentation_chart, def:presentation_basic_open}
  If \(V\) is an \(e\)-presentation chart for \(\mathcal G\) and
  \(f \in \Gamma(X, V)\), then the basic open \(D(f) \subseteq V\) is again an
  \(e\)-presentation chart.
\end{lemma}

\begin{proof}
  \leanok
  Restrict a presentation of \(\Gamma(\mathcal G, V)\) to \(D(f)\) by
  \cref{def:presentation_basic_open}.
\end{proof}

\begin{definition}[Chart fibre rank]
  \label{def:chart_fiber_rank}
  \lean{AlgebraicGeometry.Scheme.Modules.chartFiberRank}
  \leanok
  \uses{def:fiber_rank}
  For an affine open chart \(V\) and a point \(x \in V\), the \emph{chart
  fibre rank} of \(\mathcal G\) at \(x\) computed on \(V\) is the fibre rank
  of the section module \(\Gamma(\mathcal G, V)\) at the prime
  \(\mathfrak p_x \subseteq \Gamma(X, V)\) corresponding to \(x\).
\end{definition}

\begin{lemma}[Chart fibre rank under localization]
  \label{lem:chart_fiber_rank_basic_open}
  \lean{AlgebraicGeometry.Scheme.Modules.chartFiberRank_basicOpen}
  \leanok
  \uses{def:chart_fiber_rank, cor:fiber_rank_localized}
  The chart fibre rank is unchanged by passing to a basic open of the chart:
  for \(f \in \Gamma(X, V)\) and \(x \in D(f)\), the chart fibre rank computed
  on \(V\) equals that computed on \(D(f)\).
\end{lemma}

\begin{proof}
  \leanok
  \(\Gamma(\mathcal G, D(f))\) is the localization of
  \(\Gamma(\mathcal G, V)\) at \(f\), and the prime of \(x\) in
  \(\Gamma(X, D(f))\) contracts to the prime of \(x\) in \(\Gamma(X, V)\).
  Apply \cref{cor:fiber_rank_localized}.
\end{proof}

\begin{theorem}[Chart-independence of the fibre rank]
  \label{thm:chart_fiber_rank_eq}
  \lean{AlgebraicGeometry.Scheme.Modules.chartFiberRank_eq}
  \leanok
  \uses{def:chart_fiber_rank, lem:chart_fiber_rank_basic_open}
  Any two affine charts \(V, W\) containing a point \(x\) give the same chart
  fibre rank of \(\mathcal G\) at \(x\).
\end{theorem}

\begin{proof}
  \leanok
  Choose a basic open of \(V\) and a basic open of \(W\) that coincide as an
  open \(D \ni x\) contained in \(V \cap W\).  By
  \cref{lem:chart_fiber_rank_basic_open} the chart fibre rank computed on
  \(V\) and the one computed on \(W\) both equal the chart fibre rank computed
  on \(D\), hence agree.
\end{proof}

\begin{definition}[Point rank]
  \label{def:point_rank}
  \lean{AlgebraicGeometry.Scheme.Modules.pointRank}
  \leanok
  \uses{def:chart_fiber_rank, thm:chart_fiber_rank_eq}
  The \emph{point rank} of \(\mathcal G\) at a point \(x \in X\) is the
  dimension of the fibre of \(\mathcal G\) at \(x\), computed as the chart
  fibre rank on any affine chart through \(x\); it is well defined by
  \cref{thm:chart_fiber_rank_eq}.
\end{definition}

\begin{lemma}[Point rank on a chart]
  \label{lem:point_rank_eq_chart_fiber_rank}
  \lean{AlgebraicGeometry.Scheme.Modules.pointRank_eq_chartFiberRank}
  \leanok
  \uses{def:point_rank, thm:chart_fiber_rank_eq}
  For any affine chart \(V \ni x\), the point rank of \(\mathcal G\) at \(x\)
  equals the chart fibre rank computed on \(V\).
\end{lemma}

\begin{proof}
  \leanok
  Both are chart fibre ranks at \(x\); apply \cref{thm:chart_fiber_rank_eq}.
\end{proof}

\subsection{The strata ideal}

\begin{lemma}[Locality of entry-ideal membership]
  \label{lem:mem_entry_ideal_locally}
  \lean{AlgebraicGeometry.Scheme.Modules.mem_entryIdeal_of_locally}
  \leanok
  \uses{def:entry_ideal, lem:entry_ideal_basic_open}
  Let \(P = (\psi, \pi)\) be a matrix presentation of
  \(\Gamma(\mathcal G, V)\) by \(e\) generators over an affine open \(V\), and
  \(r \in \Gamma(X, V)\).  If every point of \(V\) has a basic open
  \(D(g) \ni x\) on which the restriction of \(r\) lies in the entry ideal of
  the restricted presentation \(\psi|_{D(g)}\), then \(r \in I(\psi)\).
\end{lemma}

\begin{proof}
  \leanok
  Choose for each point such a \(g\); the basic opens \(D(g)\) cover \(V\), so
  the corresponding \(g\) generate the unit ideal of \(\Gamma(X, V)\).  By
  \cref{lem:entry_ideal_basic_open}, membership of \(r|_{D(g)}\) in
  \(I(\psi|_{D(g)}) = I(\psi) \cdot \Gamma(X, D(g))\) means a power
  \(g^{n} r\) lies in \(I(\psi)\).  A section that, after multiplication by a
  power of each member of a family generating the unit ideal, lands in a
  submodule, already lies in that submodule; hence \(r \in I(\psi)\).
\end{proof}

\begin{lemma}[Chart-to-chart transfer of the entry-ideal condition]
  \label{lem:entry_ideal_chart_transfer}
  \lean{AlgebraicGeometry.Scheme.Modules.exists_basicOpen_mem_entryIdeal_of_chart}
  \leanok
  \uses{thm:entry_ideal_presentation_independent, lem:entry_ideal_basic_open}
  Let \(r \in \Gamma(X, U)\), let \(V, W \subseteq U\) be affine opens with
  \(e\)-generator presentations \(P_V, P_W\), and suppose the restriction of
  \(r\) to \(W\) lies in \(I(\psi_W)\).  Then for every point
  \(x \in V \cap W\) there is a basic open \(D(g) \subseteq V\) with
  \(x \in D(g)\) on which the restriction of \(r\) lies in
  \(I(\psi_V|_{D(g)})\).
\end{lemma}

\begin{proof}
  \leanok
  Choose a common basic open \(D \ni x\) contained in \(V \cap W\), realised
  as a basic open \(D(g)\) of \(V\) and a basic open \(D(f)\) of \(W\).  On
  \(D\) both \(\psi_V|_{D(g)}\) and \(\psi_W|_{D(f)}\) present
  \(\Gamma(\mathcal G, D)\) by \(e\) generators, so by
  \cref{thm:entry_ideal_presentation_independent} their entry ideals coincide.
  By \cref{lem:entry_ideal_basic_open} the restriction of \(r|_W\) to \(D\)
  lies in \(I(\psi_W|_{D(f)}) = I(\psi_V|_{D(g)})\).
\end{proof}

\begin{definition}[Strata ideal]
  \label{def:strata_ideal}
  \lean{AlgebraicGeometry.Scheme.Modules.strataIdeal}
  \leanok
  \uses{def:entry_ideal, def:presentation_chart}
  \source{nitsure-hilbert-quot}
  The rank-\(e\) \emph{strata ideal} of \(\mathcal G\) on an affine open \(U\)
  is the ideal of \(\Gamma(X, U)\)
  \[
    I_e(U) \;=\; \bigcap_{V \le U} \; \bigcap_{P}
      \; \bigl(\text{restriction } \Gamma(X, U) \to \Gamma(X, V)\bigr)^{-1}
        \bigl(I(\psi_P)\bigr),
  \]
  the intersection over all affine opens \(V \subseteq U\) and all matrix
  presentations \(P = (\psi_P, \pi_P)\) of \(\Gamma(\mathcal G, V)\) by \(e\)
  generators, of the preimages of the entry ideals under restriction.
\end{definition}

\begin{theorem}[Strata ideal on a chart is the entry ideal]
  \label{thm:strata_ideal_eq_entry_ideal}
  \lean{AlgebraicGeometry.Scheme.Modules.strataIdeal_eq_entryIdeal}
  \leanok
  \uses{def:strata_ideal, lem:mem_entry_ideal_locally, lem:entry_ideal_chart_transfer}
  If \(V\) is an \(e\)-presentation chart with presentation \(P\), then the
  strata ideal on \(V\) is the entry ideal of \(P\): \(I_e(V) = I(\psi_P)\).
\end{theorem}

\begin{proof}
  \leanok
  For \(\subseteq\): the pair \((V, P)\) itself occurs in the defining
  intersection, and restriction along \(V \le V\) is the identity, so every
  element of \(I_e(V)\) lies in \(I(\psi_P)\).  For \(\supseteq\): let
  \(r \in I(\psi_P)\) and let \((W, Q)\) be any presentation chart
  \(W \subseteq V\).  By \cref{lem:mem_entry_ideal_locally} it suffices to
  show that each point of \(W\) has a basic open on which \(r|_W\) lands in the
  entry ideal of the restricted \(Q\); this is
  \cref{lem:entry_ideal_chart_transfer} applied to the two charts \(W\) and
  \(V\) (using \(r|_V = r \in I(\psi_P)\)).  Hence \(r|_W \in I(\psi_Q)\) for
  every chart, i.e.\ \(r \in I_e(V)\).
\end{proof}

\begin{lemma}[Strata membership from a covering family of charts]
  \label{lem:mem_strata_ideal_of_charts}
  \lean{AlgebraicGeometry.Scheme.Modules.mem_strataIdeal_of_charts}
  \leanok
  \uses{def:strata_ideal, lem:mem_entry_ideal_locally, lem:entry_ideal_chart_transfer}
  Let \(r \in \Gamma(X, U)\) and let \((V_i)\) be a family of affine
  presentation charts \(V_i \subseteq U\), with presentations \(P_i\), whose
  underlying opens cover \(U\).  If the restriction of \(r\) to each \(V_i\)
  lies in \(I(\psi_{P_i})\), then \(r \in I_e(U)\).
\end{lemma}

\begin{proof}
  \leanok
  Let \((W, Q)\) be any presentation chart \(W \subseteq U\).  By
  \cref{lem:mem_entry_ideal_locally} it suffices to check local membership at
  each \(x \in W\); choosing an index \(i\) with \(x \in V_i\),
  \cref{lem:entry_ideal_chart_transfer} applied to \(W\) and \(V_i\) (using
  \(r|_{V_i} \in I(\psi_{P_i})\)) supplies a basic open of \(W\) on which
  \(r\) lands in the entry ideal of the restricted \(Q\).  Thus
  \(r|_W \in I(\psi_Q)\) for every chart, i.e.\ \(r \in I_e(U)\).
\end{proof}

\begin{definition}[Charts-cover hypothesis]
  \label{def:charts_cover}
  \lean{AlgebraicGeometry.Scheme.Modules.ChartsCover}
  \leanok
  \uses{def:presentation_chart}
  The module \(\mathcal G\) \emph{admits \(e\)-presentation charts} if every
  point of \(X\) lies in an \(e\)-presentation chart.  This is the standing
  hypothesis of the rank-\(e\) stratum construction; it holds on the union of
  all \(e\)-presentation charts.
\end{definition}

\begin{lemma}[Finite chart covers]
  \label{lem:charts_cover_finite}
  \lean{AlgebraicGeometry.Scheme.Modules.ChartsCover.exists_finite_charts}
  \leanok
  \uses{def:charts_cover, lem:presentation_chart_basic_open}
  If \(\mathcal G\) admits \(e\)-presentation charts, then every affine open
  \(U \subseteq X\) is covered by finitely many \(e\)-presentation charts
  contained in \(U\).
\end{lemma}

\begin{proof}
  \leanok
  Each point of \(U\) lies in some \(e\)-presentation chart; refining to a
  basic open contained in \(U\) keeps it a presentation chart
  (\cref{lem:presentation_chart_basic_open}).  These charts cover \(U\), and
  \(U\) is quasi-compact, being affine, so finitely many of them suffice.
\end{proof}

\begin{theorem}[The strata ideal is quasi-coherent]
  \label{thm:map_strata_ideal_basic_open}
  \lean{AlgebraicGeometry.Scheme.Modules.map_strataIdeal_basicOpen}
  \leanok
  \uses{def:strata_ideal, lem:charts_cover_finite, lem:mem_strata_ideal_of_charts,
    lem:entry_ideal_basic_open, thm:strata_ideal_eq_entry_ideal}
  \source{nitsure-hilbert-quot}
  Suppose \(\mathcal G\) admits \(e\)-presentation charts.  For an affine open
  \(U\) and \(f \in \Gamma(X, U)\), restriction along
  \(\Gamma(X, U) \to \Gamma(X, D(f))\) carries the strata ideal to the strata
  ideal: the image of \(I_e(U)\) generates \(I_e(D(f))\).  Consequently the
  assignment \(U \mapsto I_e(U)\) is quasi-coherent.
\end{theorem}

\begin{proof}
  \leanok
  For \(\subseteq\): any presentation chart \(W \subseteq D(f)\) is also a
  chart in \(U\), so a section of \(I_e(U)\) restricts into \(I(\psi_W)\);
  hence its restriction to \(D(f)\) lies in \(I_e(D(f))\).

  For \(\supseteq\): let \(r' \in I_e(D(f))\) and write
  \(r' = y / f^{n}\) with \(y \in \Gamma(X, U)\).  By
  \cref{lem:charts_cover_finite} choose finitely many presentation charts
  \(V_1, \dots, V_N\) covering \(U\), with presentations \(P_i\).  On each
  \(V_i\), the basic open \(D(f|_{V_i}) \subseteq D(f)\) is a chart, and the
  strata condition on \(r'\) there says \(y / f^{n}\) restricts into
  \(I(\psi_{P_i}|_{D(f|_{V_i})})\), the image ideal
  (\cref{lem:entry_ideal_basic_open}); clearing the denominator yields an
  exponent \(m_i\) with \(f^{m_i} y|_{V_i} \in I(\psi_{P_i})\).  Setting
  \(m = \max_i m_i\), multiplying up gives
  \(f^{m} y|_{V_i} \in I(\psi_{P_i})\) for all \(i\), so
  \(f^{m} y \in I_e(U)\) by \cref{lem:mem_strata_ideal_of_charts}.  Therefore
  \(r' = (f^{m} y) / f^{m + n}\) lies in the image of \(I_e(U)\) under
  restriction.  (\cref{thm:strata_ideal_eq_entry_ideal} identifies the strata
  ideal with the entry ideal on each chart, licensing the denominator
  clearing.)
\end{proof}

\subsection{The rank stratum}

\begin{definition}[Strata ideal-sheaf datum]
  \label{def:strata_data}
  \lean{AlgebraicGeometry.Scheme.Modules.strataData}
  \leanok
  \uses{def:strata_ideal, thm:map_strata_ideal_basic_open}
  Suppose \(\mathcal G\) admits \(e\)-presentation charts.  The rank-\(e\)
  \emph{strata ideal-sheaf datum} on \(X\) is the assignment
  \(U \mapsto I_e(U)\) of \cref{def:strata_ideal}, which is a quasi-coherent
  ideal sheaf by \cref{thm:map_strata_ideal_basic_open}.
\end{definition}

\begin{definition}[Rank-\(e\) stratum]
  \label{def:stratum}
  \lean{AlgebraicGeometry.Scheme.Modules.stratum}
  \leanok
  \uses{def:strata_data}
  \source{nitsure-hilbert-quot}
  The rank-\(e\) \emph{stratum} of \(\mathcal G\) is the closed subscheme of
  \(X\) cut out by the strata ideal-sheaf datum of \cref{def:strata_data}.
\end{definition}

\begin{definition}[Closed immersion of the stratum]
  \label{def:stratum_immersion}
  \lean{AlgebraicGeometry.Scheme.Modules.stratumι}
  \leanok
  \uses{def:stratum}
  The canonical morphism from the rank-\(e\) stratum to \(X\) is a closed
  immersion.
\end{definition}

\begin{theorem}[Support of the rank-\(e\) stratum]
  \label{thm:mem_range_stratum}
  \lean{AlgebraicGeometry.Scheme.Modules.mem_range_stratumι_iff}
  \leanok
  \uses{def:stratum_immersion, thm:strata_ideal_eq_entry_ideal, thm:entry_ideal_prime_iff,
    def:point_rank, lem:point_rank_eq_chart_fiber_rank}
  \source{nitsure-hilbert-quot}
  A point \(x \in X\) lies in the image of the closed immersion of the
  rank-\(e\) stratum if and only if the point rank of \(\mathcal G\) at \(x\)
  equals \(e\); that is, the support of the stratum is exactly the locus where
  the fibre of \(\mathcal G\) has dimension exactly \(e\).
\end{theorem}

\begin{proof}
  \leanok
  Choose an \(e\)-presentation chart \(V \ni x\) with presentation \(P\).  The
  image of the closed immersion is the support of the ideal sheaf, which meets
  \(V\) in the zero locus of \(I_e(V)\); and \(I_e(V) = I(\psi_P)\) by
  \cref{thm:strata_ideal_eq_entry_ideal}.  A point of an affine open lies in
  the zero locus of an ideal of sections iff its prime contains that ideal, so
  \(x\) lies in the stratum iff \(I(\psi_P) \subseteq \mathfrak p_x\).  By
  \cref{thm:entry_ideal_prime_iff} this holds iff the fibre rank of
  \(\Gamma(\mathcal G, V)\) at \(\mathfrak p_x\) --- the chart fibre rank at
  \(x\), equal to the point rank by
  \cref{lem:point_rank_eq_chart_fiber_rank} --- equals \(e\).
\end{proof}

\section{Universal property of the strata}
\label{sec:flatstrat_universal}

\begin{theorem}

\leanok
[Universal property of the flat-locus stratification (case \(n = 0\))]
  \label{thm:flattening_stratification_universal}
  \lean{AlgebraicGeometry.flatLocusStratification_universal}
  \uses{def:coherent_sheaf_flat}
  \source{nitsure-hilbert-quot}
  Let \(S\) be a noetherian scheme and \(\F\) a coherent \(\OO_S\)-module.
  Then there exist a finite index set \(I\) and locally-closed immersions
  \(\iota_f : S_f \to S\) (\(f \in I\)) with pairwise disjoint images
  set-theoretically covering \(|S|\), such that the pullback of \(\F\)
  along \(\iota_f\) is flat over \(S_f\) for every \(f\) (in the sense of
  \cref{def:coherent_sheaf_flat}), and such that, writing
  \(S' := \coprod_{f \in I} S_f\) and \(i : S' \to S\) for the morphism
  assembled from the \(\iota_f\), the pair \((S', i)\) has the universal
  property: every morphism \(\varphi : T \to S\) for which
  \(\varphi^* \F\) is flat over \(T\) factors \emph{uniquely} through
  \(i\).

  This is part (ii) of Nitsure's flattening-stratification theorem in the
  special case \(n = 0\), stated with flatness in place of local freeness of
  constant rank: for a finitely presented module the two notions agree
  locally, the rank decomposing \(T\) into clopen pieces.  Nitsure states an
  ``if and only if''; the converse direction --- any morphism factoring
  through \(i\) has flat pullback --- is stability of flatness under base
  change and is recorded separately once needed.  The universal property for
  a relative family (Nitsure's part (ii) for \(\F\) on \(\PP^n_S\), hence
  for projective \(\pi : X \to S\)) additionally requires the
  cohomology-and-base-change layer of \cref{lem:noetherian_induction_strata}
  and is not stated separately here; for a general \emph{proper} family the
  universal property is not asserted by any source in the reference library,
  and the reduced strata of \cref{lem:flat_locus_reduction_lean} do not
  witness it (over \(S = \Spec k[\varepsilon]\) with \(\F = \OO_S\) flat,
  the identity must factor through the stratification, forcing the canonical
  non-reduced stratum structure).
\end{theorem}

\begin{proof}
  \uses{lem:flat_locus_open, def:stratum, def:stratum_immersion,
    thm:mem_range_stratum, thm:map_strata_ideal_basic_open}
  By the case \(n = 0\) of the flattening stratification
  (\cref{lem:flat_locus_open}), with its canonical scheme structure: locally
  on \(S\), a basis of \(\F \otimes \kappa(s)\) prolongs by Nakayama to a
  right-exact presentation
  \(\OO_V^{\oplus m} \xrightarrow{\psi} \OO_V^{\oplus e} \to \F|_V \to 0\)
  on a neighbourhood \(V\) of \(s\), and the closed subscheme
  \(V_e \subseteq V\) cut out by the ideal of matrix entries of \(\psi\)
  represents the locus where \(\F\) becomes locally free of rank \(e\):
  for \(f : T \to V\) the pulled-back presentation stays right-exact, so
  \(f^*\F\) is locally free of rank \(e\) if and only if \(f^*\psi = 0\),
  if and only if \(f\) factors through \(V_e\).  The local strata glue by
  this representing property into locally-closed subschemes
  \(S_e \subseteq S\) with \(|S_e|\) the rank-\(e\) locus; the family is
  finite because the fibre rank of a coherent module is bounded on a
  noetherian scheme, and the images are disjoint and cover \(|S|\) since
  every point has a well-defined fibre rank.

  For the universal property, let \(\varphi : T \to S\) have
  \(\varphi^*\F\) flat over \(T\).  A finitely presented flat module is
  locally free, and its rank is locally constant, so \(T\) decomposes into
  clopen pieces \(T_e\) on which \(\varphi^*\F\) is locally free of
  constant rank \(e\); by the representing property each restriction
  \(\varphi|_{T_e}\) factors uniquely through \(S_e\), and the
  factorizations assemble to \(\varphi = i \circ \tilde\varphi\) with
  \(\tilde\varphi : T \to \coprod_e S_e\).  Uniqueness: the strata have
  disjoint images, so any factorization through \(i\) induces the same
  decomposition of \(T\) into clopen preimages, and on each piece the
  component through \(S_e\) is unique because \(\iota_e\) is an immersion,
  hence a monomorphism.
\end{proof}

\section{Application: relative curves}
\label{sec:flatstrat_curve_application}

In the Route~A consumer setting, the input to the flattening-stratification
machinery is the pullback \(C \times_k T \to T\) where \(C\) is a smooth proper
curve over a field \(k\) and \(T\) is any \(k\)-scheme.  The hypotheses of
Nitsure's \cref{thm:flattening_stratification_exists} demand
\emph{projective} relative dimension, not merely \emph{proper}; we record that
no extra work is required to instantiate the theorem in our setting.

\begin{lemma}
[Smooth proper curve over \(k\) is projective]
  \label{lem:smooth_proper_curve_projective}
  \lean{AlgebraicGeometry.TODO.smoothProperCurveProjective}
  % SOURCE: [Hartshorne], II.6.7 / II.7.6 (curve \(\Rightarrow\) quasi-projective;
  % proper \(\Rightarrow\) projective for curves), and standard consequences in
  % chapter III. The precise statement ``every smooth proper curve over a field
  % is projective'' is one of the classical equivalences for \(\dim 1\) and is
  % textbook material; this block is project-bespoke (no verbatim source
  % quote is reproduced here as the result is folklore-level and cited only
  % structurally for the Route~A reduction).
  \textit{Source: classical; cf.\ [Hartshorne], II.6.7 \& II.7.6 (curves).}
  Let \(C\) be a smooth proper curve over a field \(k\).  Then \(C \to \Spec k\) is
  projective.  Consequently, for any noetherian \(k\)-scheme \(T\), the relative
  curve \(C \times_k T \to T\) is projective in the sense of Grothendieck (there
  exists a relatively very ample line bundle on \(C \times_k T\) over \(T\)).
\end{lemma}

\begin{proof}
  Projectivity of a smooth proper curve over a field is classical
  ([Hartshorne], II.6.7 / II.7.6; alternatively: every divisor of large
  positive degree on a smooth proper curve \(C\) is very ample, providing the
  embedding \(C \hookrightarrow \PP^N_k\) for some \(N\)). Pulling the ample line
  bundle back along \(C \times_k T \to C\) produces a relatively very ample line
  bundle on \(C \times_k T\) over \(T\).
\end{proof}

\begin{corollary}

\leanok
[Flattening stratification for a coherent sheaf on a relative curve]
  \label{cor:flattening_stratification_curve}
  \lean{AlgebraicGeometry.flatteningStratification.ofCurve}
  \uses{lem:smooth_proper_curve_projective, thm:flattening_stratification_exists}
  Let \(k\) be a field, \(C\) a smooth proper curve over \(k\), and \(T\) a noetherian
  \(k\)-scheme.  Let \(\F\) be a coherent sheaf on \(C \times_k T\). Then the
  flattening-stratification conclusion of
  \cref{thm:flattening_stratification_exists} applies: there is a finite
  set \(I\) of Hilbert polynomials and locally-closed subschemes
  \(T_f \hookrightarrow T\) for \(f \in I\), set-theoretically covering \(T\)
  disjointly, such that (a) \(\F|_{C \times_k T_f}\) is flat over \(T_f\) for
  each \(f\), and (b) any morphism \(\varphi : T' \to T\) such that
  \(\varphi^* \F\) on \(C \times_k T'\) is \(T'\)-flat factors uniquely through some
  \(T_f\).
\end{corollary}

\begin{proof}
  
  By \cref{lem:smooth_proper_curve_projective}, \(C \times_k T \to T\) is
  projective, hence factors through a closed immersion into \(\PP^n_T\) for some
  \(n \ge 0\).  Push \(\F\) forward along the closed immersion (extending by zero
  outside the image) to a coherent sheaf \(\tilde \F\) on \(\PP^n_T\); this
  operation preserves \(T\)-flatness in both directions because the closed
  immersion is a finite morphism.  Apply
  \cref{thm:flattening_stratification_exists} to \(\tilde \F\) on
  \(\PP^n_T\) to obtain \(\{T_f\}\) with the stated universal property.
\end{proof}

\section{Mathlib status}
\label{sec:flatstrat_mathlib_status}

\emph{Formalization status (run 0010, T12 r7).}  The Lean-typed existence
statement (\texttt{AlgebraicGeometry.flatteningStratification}: finite
disjoint locally-closed cover with flat pullbacks, without the
Hilbert-polynomial indexing) is proved sorry-free and axiom-clean, through
\cref{lem:flat_locus_reduction_lean} and the transfer lemmas of
\cref{subsec:flat_transfer_strata}; the Lean sections \S3--\S6 of the
statements live in \texttt{GenericFlatnessGeometric.lean} because the proof
consumes generic flatness, whose glue needs the QuotScheme engine.  The
\(n = 0\) stratification \texttt{flatLocusStratification}
(\cref{lem:flat_locus_stratification_lean}) is likewise proved axiom-clean,
derived from the existence theorem at \(\pi = \mathrm{id}_S\) through the
iso-insensitivity of the flatness predicate
(\texttt{coherentSheafFlat\_of\_comp\_isIso}) and empty-scheme padding of
the index set.  The remaining Lean gaps in this chapter are the \(n = 0\)
universal property (\texttt{flatLocusStratification\_universal},
\cref{thm:flattening_stratification_universal}, the canonical entry-ideal
strata) and the Hilbert-polynomial-indexed refinement of the main statement,
which additionally awaits the coherent-\(\chi\) substrate.

Mathlib (master \texttt{b80f227}) provides:
\begin{itemize}
  \item \texttt{Module.Flat} (\texttt{Mathlib.RingTheory.Flat.Basic}) --- the
    module-theoretic notion of flatness, applicable affine-locally to the
    stalks \(\F_x\).
  \item \texttt{Mathlib.AlgebraicGeometry.Morphisms.Flat} ---
    \texttt{IsFlat} as a morphism property and basic stability lemmas; this is
    flatness of the structure morphism, \emph{not} flatness of a quasi-coherent
    module over the base.
  \item \texttt{Module.FaithfullyFlat} and the descent infrastructure that
    consumes it.
\end{itemize}

\noindent What Mathlib does \emph{not} yet have, and that this chapter will
introduce in the new file
\texttt{AlgebraicJacobian/Picard/FlatteningStratification.lean}:
\begin{itemize}
  \item A predicate \texttt{AlgebraicGeometry.Scheme.CoherentSheafFlat}
    capturing \(S\)-flatness of a coherent \(\OO_X\)-module on a finite-type
    \(S\)-scheme (\cref{def:coherent_sheaf_flat}). This is the
    sheaf-level lift of \texttt{Module.Flat} via the affine-local
    characterization in the second paragraph of the definition.
  \item The constructive existence proof
    \texttt{AlgebraicGeometry.flatteningStratification}
    (\cref{thm:flattening_stratification_exists}) from Nitsure~\S4, proved
    through \texttt{flatLocusReduction}
    (\cref{lem:flat_locus_reduction_lean}, the Noetherian-induction
    reduction via generic flatness) and packaged with injective labels by
    \texttt{flatLocusAssembly} (\cref{lem:flat_locus_assembly_lean}); the
    \(n = 0\) case \texttt{flatLocusStratification}
    (\cref{lem:flat_locus_stratification_lean}) is derived back from the
    existence theorem at \(\pi = \mathrm{id}_S\).
  \item The universal-property packaging
    \texttt{AlgebraicGeometry.flatLocusStratification\_universal}
    (\cref{thm:flattening_stratification_universal}), bundling the \(n = 0\)
    stratification with the unique factorisation of morphisms with flat
    pullback through the stratum coproduct \(\coprod_f S_f \to S\); the
    projective-family upgrade needs \(\PP^n_S\) vocabulary and
    cohomology-and-base-change, and an eventual
    \texttt{CategoryTheory.Functor.RepresentableBy}-style refinement remains
    future work.
  \item The dévissage layer of the algebraic generic-freeness lemma
    (\cref{sec:generic_freeness_devissage}) is formalized and proved
    (splicing, torsion base case, transport, prime-filtration assembly
    \texttt{GenericFreeness.genericFlatnessAlgebraic}); the sole remaining
    algebraic gap is the normalisation core
    (\cref{lem:generically_free_domain_core}).
  \item The relative-curve specialisation
    \texttt{AlgebraicGeometry.flatteningStratification.ofCurve}
    (\cref{cor:flattening_stratification_curve}) for the Route~A
    consumer.
\end{itemize}

Inputs from Mathlib that the proof consumes:
\begin{itemize}
  \item \texttt{CategoryTheory.Limits.Cofan} / \texttt{Sigma} for the coproduct
    \(\coprod_f S_f\).
  \item \texttt{AlgebraicGeometry.IsLocallyClosedImmersion} (and the locally-closed
    subscheme API) for the strata \(S_f \hookrightarrow S\).
  \item \texttt{IsNoetherian} / \texttt{Mathlib.RingTheory.Noetherian} for the
    Noetherian-induction termination argument.
  \item \texttt{AlgebraicGeometry.IsProjective} (or
    \texttt{ProjectiveMorphism}) and the relative-\(\PP^n\) construction, to
    state and unpack the projective-morphism hypothesis on
    \(\pi : \PP^n_S \to S\).
  \item Higher-direct-image base-change lemmas (\texttt{Mathlib.AlgebraicGeometry
    .Cohomology}, or a Stacks-tag-02KH-shaped lemma; cf.\ the existing
    \cref{chap:Picard_RelativeSpec} use of the same tag) for the \(E_i = \pi_*\F(N+i)\)
    construction in \cref{lem:noetherian_induction_strata}.
  \item Castelnuovo--Mumford regularity / Serre vanishing for the
    \texttt{N\_1}-existence step.  Mathlib does not yet have these in the
    relative form needed; a future sub-lemma may stub them or rely on
    \texttt{Module.IsTorsion} arguments on the graded module
    \(\bigoplus_{m \ge N} \pi_* \F(m)\).
\end{itemize}

\section{Out of scope}
\label{sec:flatstrat_out_of_scope}

This chapter packages \emph{only} the flattening stratification of a coherent
sheaf on a projective morphism, plus the relative-curve specialisation. The
following downstream / parallel constructions live in separate sibling
chapters and are explicitly out of scope here:
\begin{itemize}
  \item \textbf{Castelnuovo--Mumford regularity} (Nitsure~\S2) --- input to the
    Quot construction, used implicitly inside the proof of
    \cref{lem:noetherian_induction_strata} as the assertion that the
    Hilbert function stabilises beyond a uniform \(N\).  Its own blueprint home
    is the planned chapter \texttt{Picard\_CastelnuovoMumford.tex} (not yet
    scaffolded).
  \item \textbf{Quot scheme construction} (Nitsure~\S5, A.2.b) --- consumer of
    the present chapter; its home is the planned \texttt{Picard\_QuotScheme.tex}.
  \item \textbf{Relative Picard functor} (A.1.c) ---
    \texttt{Picard\_RelPicFunctor.tex} (planned).
  \item \textbf{Picard scheme representability for a smooth proper curve}
    (A.2.c) --- the assembly theorem fed by both the Quot construction (which
    consumes this chapter) and the relative Picard functor; home is the
    planned \texttt{Picard\_FGAPicRepresentability.tex}.
\end{itemize}
